---
title: "bda-politeRSA-S-rwebppl"
author: "M. H. Tessler, Erica Yoon"
date: "1/9/2017"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(binom)
library(rwebppl)
library(jsonlite)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
library(coda)
library(magrittr)
library(ggthemes)
library(forcats)
# set path to working dir
 # local.path <- "~/Documents/research/polgrice/"
local.path <- "~/Documents/Research/polgrice_GIT/"
#source(paste(local.path, "experiment/data_analysis/markdown/polgrice_S.R", sep=""))

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
# HPDhi<- function(s){
hdi_upper <- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
# HPDlo<- function(s){
hdi_lower <- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   

```

# Literal semantics data

```{r eval=FALSE}
d.lit <- read.csv(paste(local.path,
                        "experiment/data_analysis/data/literalSemantics_wNeg.csv",
                        sep = ""))
```

Summarize behavioral data (in terms of number of "yes"es)

```{r eval=FALSE}
d.lit.summary <- d.lit %>%
  group_by(state, utterance) %>%
  summarize(k = sum(judgment),
            n = n())
```

## Model

```{r eval=FALSE}
literalSemanticsModel <- '
var n = data[0]["n"]
var k = data[0]["k"]

var literalSemantics = function(){
	var theta = uniform(0,1)
  observe( Binomial({n: n, p: theta}), k )
  return {theta: theta}
}
'
```

Run model, for all sites, states, utterances

```{r eval=FALSE}
# sites <- levels(d.lit.summary$site)
states <- levels(factor(d.lit.summary$state))
utterances <- levels(d.lit.summary$utterance)

litSemantics.results <- data.frame()

for (st in states){
  print(st)
  for (utt in utterances){
        print(utt)

    d.lit.pass <- d.lit.summary %>% 
      filter((state == st) & (utterance == utt))
    
    rs <- webppl(literalSemanticsModel,
                 data = d.lit.pass,
                 data_var = "data",
                 inference_opts = list(
                   method = "rejection",
                   samples = 10000
                 ),
                 model_var = "literalSemantics",
                 output_format = "samples")
    
    rs.summary <- rs %>%
      rename(theta = value.theta) %>%
      # summarize(MAP = estimate_mode(value.theta),
      #           cred_low = hdi_lower(value.theta),
      #           cred_upper = hdi_upper(value.theta)) %>%
      mutate(state = st, utterance = utt)
    
    litSemantics.results <- bind_rows(litSemantics.results,
                                      rs.summary)
  }
}

# write.csv(litSemantics.results,paste(local.path,
#                          "model/data/literal_semantics_wNeg_10ksamples_fullPosterior.csv",sep=""), # CHANGE FILE NAME AS NEEDED
#           row.names=FALSE)
```

Binomial is conjugate to Beta, we can solve for the posterior analytically

```{r}

d.lit.summary <- d.lit.summary %>%
  mutate(posterior_b1 = 1 + k,
         posterior_b2 = 1 + n - k, 
         MAP = (posterior_b1 - 1) / (posterior_b1 + posterior_b2 - 2),
          cred_low  = qbeta(.025, posterior_b1, posterior_b2),
         cred_upper = qbeta(.975, posterior_b1, posterior_b2))

litSemantics.results.org <- d.lit.summary %>%
    mutate(positivity = factor(as.numeric(grepl("yes", utterance)), 
                             levels = c(0, 1), 
                             labels = c("negative","positive"))) %>%
  mutate(utterance = substring(utterance, 5)) %>%
  mutate(utterance = ordered(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")))

ggplot(data=litSemantics.results.org, 
       aes(x=state, y=MAP, col=utterance, group=utterance)) +
  geom_line() +
  facet_grid(positivity~utterance) +
  geom_errorbar(aes(ymin=cred_low,ymax=cred_upper, width=.1))

```

# production prediction data

```{r}
d.utterance <- read.csv(paste(
  local.path, 
  "experiment/data_analysis/data/speaker.csv"
  , sep="")
)  %>%
    filter(utterance != "NA_NA")
```

```{r eval=FALSE}

d <- d.utterance %>%
  separate(utterance, into = c("positivity", "utterance"), sep = "_") %>%
  mutate(true_state = as.factor(true_state),
         goal = as.factor(goal),
         positivity = as.factor(positivity),
         utterance = as.factor(utterance)
          )


ms2 <- d %>%
  filter(!is.na(positivity), !is.na(utterance)) %>% # why is there NA?
  group_by(true_state, goal) %>%
  summarise(n.total=n())

ms3 <- d %>%
  filter(!is.na(positivity), !is.na(utterance)) %>% # why is there NA?
  group_by(true_state, goal, positivity, utterance) %>%
  summarize(n = n())

ms <- left_join(ms2, ms3) %>%
  group_by(true_state, goal, positivity, utterance) %>%
  summarize(mean = n / n.total,
            ci_lower = binom.bayes(n, n.total)$lower,
            ci_upper = binom.bayes(n, n.total)$upper) 
  
ms_fake <- expand.grid(true_state=levels(ms$true_state),
                             goal=levels(ms$goal),
                             positivity=levels(ms$positivity), 
                             utterance=levels(ms$utterance))

ms.all <- left_join(data.frame(ms_fake),data.frame(ms))

ms.all[is.na(ms.all)] <- 0
levels(ms.all$true_state) <- c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts")
levels(ms.all$goal) <- c("want both", "want to be informative", "want to make listener feel good")
levels(ms.all$positivity) <- c("negation", "no negation")

ms.all %<>%
  mutate(utterance = fct_relevel(utterance, "terrible", "bad", "okay", "good", "amazing"),
         goal = fct_relevel(goal, "want to be informative", "want to make listener feel good", "want both"))

ggplot(data=ms.all, aes(x=utterance, y=mean, fill=positivity, group = positivity, color = positivity)) +
  #geom_bar(stat="identity", position=position_dodge()) +
  geom_line()+
  facet_grid(goal~true_state) +
  xlab("no negation (It was ~) vs negation (It wasn't ~) ") +
  ylab("proportion chosen") +
  # ggtitle("What would the speaker say given their goals?") +
  geom_linerange(aes(ymin=ci_lower,ymax=ci_upper), position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  #theme_bw()+
  ylim(0,1)+
  scale_color_solarized()+
  ggthemes::theme_few()+
  #theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))

#ggsave("~/Desktop/politeness.pdf", width = 10)
```


```{r}
with(filter(d, goal == "informative"),
  table(positivity, utterance, true_state))
```

```{r}
with(filter(d, goal == "social"),
  table(positivity, utterance, true_state))
```

```{r}
with(filter(d, goal == "both"),
  table(positivity, utterance, true_state))
```

```{r}
with(filter(d, goal == "informative"),
  table(positivity, true_state))
```

```{r}
with(filter(d, goal == "social"),
  table(positivity, true_state))

```

```{r}
with(filter(d, goal == "both"),
  table(positivity, true_state))
```


Polite RSA + data analysis model

```{r}
fullModel <- '
var utterances = ["yes_terrible","yes_bad","yes_okay","yes_good","yes_amazing",
                  "not_terrible","not_bad","not_okay","not_good","not_amazing"
                  ];

var states = [1,2,3,4,5];

var statePrior = function(){
  return uniformDraw(states);
};

var isNegation = function(utt){
  return (utt.split("_")[0] == "not")
};

// measured in Experiment 1
var literalSemantics = dataFromR.literalSemantics;

 var litSemanticsPosteriorObj = _.object(map(function(u){
  return [u, _.object(map(function(s){
      var litParams = _.where(literalSemantics, {state: s, utterance: u})[0]
    return [s, {a: litParams.posterior_b1, b: litParams.posterior_b2}]
 }, states))]
 }, utterances))

// e.g. {"amazing": { 1: [w1, w2, ... , wn], 2: [ ... ], ... }, "terrible": {1: [, ...]} }
var round = function(x){
	return Math.round(x * 100) / 100
}

var weightBins = map(round, _.range(0.0001, 1, 0.05))
var phiWeights = repeat(weightBins.length, function(){1})
  
var goalWeightPrior = Infer({method: "enumerate"}, function(){
  return uniformDraw(weightBins);
})

var nBins = weightBins.length;
var kernelWidth = nBins / 4;

var kernelFn = function(prevVal){
  var i = weightBins.indexOf(prevVal);
  var upper = (i + kernelWidth) > nBins ? nBins : i + kernelWidth;
  var lower = (i - kernelWidth) < 0 ? 0 : i - kernelWidth;
  return Categorical({vs: weightBins.slice(lower,upper), 
                      ps: phiWeights.slice(lower,upper)})
}


var data = dataFromR.data;

var goals = _.uniq(_.pluck(data, "goal"));

var states = _.uniq(_.pluck(data, "true_state"));
// var utterances = _.uniq(_.pluck(data, "utterance"));


var dataAnalysis = function(){
    
    var litSemantics = _.object(map(function(u){
    	return [u, _.object(map(function(s){
        var litParams = litSemanticsPosteriorObj[u][s];
         return [s, beta(litParams)]
      }, states))]
     }, utterances))

    var RSAparameters = {
      speakerOptimality: uniformDrift({a: 0, b: 20, width:2}),
      speakerOptimality2: uniformDrift({a:0, b: 5, width: 0.5}),
      alpha: uniformDrift({a: 0, b: 5, width:0.5}),
      cost: uniformDrift({a: 1, b: 10, width:0.25})
    };

    var cost_yes = 1;
    var uttCosts = map(function(u) {return isNegation(u) ? Math.exp(-RSAparameters.cost) : Math.exp(-cost_yes)},
            utterances)
    
    var utterancePrior = Infer({method: "enumerate"}, function(){
      return  utterances[discrete(uttCosts)];
    });

    var meaning = function(words, state){
       //display(litSem[words][state])
      return flip(litSemantics[words][state]);
    }; 
    
    var listener0 = cache(function(utterance) {
      Infer({method: "enumerate"}, function(){
        var state = uniformDraw(states);
        var m = meaning(utterance, state);
        condition(m);
        return state;
      });
    }, 10000);
    
    var speaker1 = cache(function(state, speakerGoals) {
      Infer({method: "enumerate"}, function(){
        var utterance = sample(utterancePrior);
    
        var speakerOptimality = RSAparameters.speakerOptimality;
        var alpha = RSAparameters.alpha;
    
        var L0 = listener0(utterance);
        
        var epistemicUtility = L0.score(state);
        var socialUtility = expectation(L0, function(s){return alpha*s});
        
        var eUtility = speakerGoals.phi*epistemicUtility;
        var sUtility = (1-speakerGoals.phi)*socialUtility;
    
        var speakerUtility = eUtility+sUtility;
    
        factor(speakerOptimality*speakerUtility);
        
        return utterance;
      })
    }, 10000)
    
    var listener1 = cache(function(utterance) {
     Infer({method: "enumerate"}, function(){
    
       var speakerGoals = {
         phi:categorical ({vs: weightBins, ps: phiWeights})
       }
       
       var state = uniformDraw(states);
    
       var S1 = speaker1(state, speakerGoals)
       observe(S1, utterance)
        
       return {
         state: state,
         goals: speakerGoals
       }
        })
    }, 10000)
    
    var speaker2 = cache(function(exptCondInfo) {
     Enumerate(function(){
       var state = exptCondInfo.state;
       var intendedGoals = exptCondInfo.goalWeights;
       var utterance = sample(utterancePrior);
    
       var L1 = listener1(utterance)
    
       factor(RSAparameters.speakerOptimality2 * L1.score({"state":state, "goals":intendedGoals}))

       return utterance
    
     })
    }, 10000)



    var goalWeightsAndPostPred = map(function(goal){
      
      var goalWeights = {
        //phi: uniformDraw(weightBins)
        phi: sample(goalWeightPrior, {driftKernel: kernelFn})
        //phi: uniformDrift({a:0, b:1, width:0.2})
      }

    var postPred = map(function(state){
      
     var utteranceData = _.pluck(_.where(data, {true_state: state, goal: goal}), "utterance"); 

      var exptConditionInfo = {
        state: state, 
        utterance: false,
        goalWeights: goalWeights
      };


      var RSApredictions = speaker2(exptConditionInfo);
      //var RSApredictions = speaker1(exptConditionInfo.state, exptConditionInfo.goalWeights);


      mapData({data: utteranceData}, 
          function(d){ 
               // display(RSApredictions.score(d))
              observe(RSApredictions, d) 
          });

      var postSupport = RSApredictions.support(); // all utterances in the posterior

      var postPredictive = map(function(u){
        return {
          key: "posteriorPredictive",
          goal: goal,
          state: state,
          utt: u,
          val: Math.exp(RSApredictions.score(u))
        }
      }, postSupport)

      var negative_utterances = [
        "not_terrible","not_bad","not_okay","not_good","not_amazing"
      ]

      var negEndorsement = sum(map(function(u){
          return Math.exp(RSApredictions.score(u))
      }, negative_utterances))
      
      return _.flatten([postPredictive, {
          key: "posteriorPredictive",
          goal: goal,
          state: state,
          utt: "negation",
          val: negEndorsement
        }])
      
    }, states)

    return [postPred, 
            {key: "phi", goal: goal, state: "NA", utt: "NA", val: goalWeights.phi},
            //{key: "weightHonest", goal: goal, state: "NA", utt: "NA", val: goalWeights.honesty},
            //{key: "weightKind", goal: goal, state: "NA", utt: "NA", val: goalWeights.kindness}
            ]
    
  }, goals)

  var returnList = _.flatten([goalWeightsAndPostPred, 
                              //litSemantics,
                              {key: "speakerOptimality", goal: "NA", utt: "NA", state: "NA", val: RSAparameters.speakerOptimality},
                              {key: "speakerOptimality2", goal: "NA", utt: "NA", state: "NA", val: RSAparameters.speakerOptimality2},
                              {key: "alpha", goal: "NA", utt: "NA", state: "NA", val: RSAparameters.alpha},
                              {key: "cost", goal: "NA", utt: "NA", state: "NA", val: RSAparameters.cost}
                              ])
  
  var returnObj = _.object(map(function(i){
    [i.key + "_" + i.goal + "_" + i.state + "_" + i.utt, i.val]
  }, returnList))

  return returnObj
    
}

'
```


Run Full BDA model

```{r}
bda.utterance.results <- data.frame()

litSemantics.toPass <- d.lit.summary %>% select(-k, -n)
dataToWebPPL <- list(literalSemantics = litSemantics.toPass,
                     data = d.utterance)

numSamples <- 40000
nChains <- 3
bda.utterance.results <- webppl(fullModel,
                            data = dataToWebPPL,
                            data_var = "dataFromR",
                            inference_opts = list(method = "MCMC", 
                                                #  kernel = list(HMC = list(steps = 5, stepSize = 0.001)),
                                                  samples = numSamples,
                                                  burn = numSamples/2,
                                                  verbose = TRUE),
                            model_var = "dataAnalysis",
                            output_format = "samples",
                            chains = nChains,
                            cores = nChains)

write.csv(bda.utterance.results,paste(local.path,
            "model/results/bda-politeRSA-speaker2GivenStateGoal_40000burn20000-3chains.csv",sep=""), # CHANGE FILE NAME AS NEEDED
row.names=FALSE)


```

›
# exp vs. BDA analysis

```{r callOnData}
# bda.utterance.results <- read.csv(paste(local.path,
#                         "model/results/bda-politeRSA-speaker1GivenStateGoal-80000burn40000.csv",
#                         sep = ""))
# bda.utterance.results <- fread(paste(local.path,
#                         "model/results/bda-politeRSA-speaker1GivenStateGoal_10000burn5000_test2.csv",
#                         sep = ""))
# 
# summary(bda.utterance.results$speakerOptimality_NA_NA_NA)
# summary(bda.utterance.results$alpha_NA_NA_NA)
# 

bda.utterance.tidy <- bda.utterance.results %>%
    select(contains("posteriorPredictive"), -contains("negation")) %>%
    gather(key, val) %>%
   #  mutate(key = gsub("value.", "", key)) %>% # for mht
    separate(key, into = c("param", "goal", "state", "positivity", "utterance")) %>%
  mutate(utterance = factor(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")),
         positivity = factor(positivity, labels = c("neg", "no_neg")),
         positivity = factor(positivity, levels = c("no_neg", "neg"))) %>%
    group_by(goal, state, positivity, utterance) %>%
  filter(!is.na(val)) %>%
  # summarise(val = mean(val))
  summarize(MAP = estimate_mode(val),
            ci_lower = hdi_upper(val),
            ci_upper = hdi_lower(val)) %>% ungroup()

  # multi_boot_standard(column = "val") %>%
  # mutate(val = mean)

# write.csv(bda.state.tidy,paste(local.path,
#                       "model/results/bda-politeRSA-speaker1GivenStateGoal-mixtureWeight-postLongForm.csv",sep=""), # CHANGE FILE NAME AS NEEDED
#           row.names=FALSE)

#bda.utterance.tidy <- fread(paste(local.path,
#                        "model/results/bda-politeRSA-speaker1GivenStateGoal-mixtureWeight-postLongForm.csv",
#                        sep = ""))
```

### Parameters

#### optimalities, cost, value-scaling

```{r alphaOptimalityCost}

bda.utterance.results %>%
   # select(value.speakerOptimality_NA_NA_NA,value.speakerOptimality2_NA_NA_NA, value.alpha_NA_NA_NA, value.cost_NA_NA_NA) %>%
  select(speakerOptimality_NA_NA_NA, speakerOptimality2_NA_NA_NA, alpha_NA_NA_NA, cost_NA_NA_NA) %>%
    gather(key, val) %>%
     # mutate(key = gsub("value.", "", key)) %>% # for mht
    separate(key, into = c("param", "goal", "state", "utterance")) %>%
  ggplot(., aes(x = val))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")

bda.utterance.results %>% 
   # select(value.speakerOptimality_NA_NA_NA,value.speakerOptimality2_NA_NA_NA, value.alpha_NA_NA_NA, value.cost_NA_NA_NA) %>%
  select(speakerOptimality_NA_NA_NA, speakerOptimality2_NA_NA_NA, alpha_NA_NA_NA, cost_NA_NA_NA) %>%
    gather(key, val) %>% 
     # mutate(key = gsub("value.", "", key)) %>% # for mht
    separate(key, into = c("param", "goal", "state", "utterance")) %>%
  group_by(param) %>%
  summarize(
    MAP = estimate_mode(val),
            ci_lower = hdi_lower(val),
            ci_upper = hdi_upper(val))

# ggsave("~/Desktop/politeModelParamS2.pdf", width = 10)

  
```

#### goal weights

```{r inferredGoalWeight}
bda.state.tidy2 <- bda.utterance.results %>% 
    select(contains("phi")) %>%
    # select(contains("weight")) %>%
    gather(key, val) %>% 
    # mutate(key = gsub("value.", "", key)) %>% # for mht
    separate(key, into = c("param", "goal", "state", "utterance")) %>%
  mutate(epistemic = val,
         social = 1-val) %>%
  select(-val) %>%
  gather(weight, val, epistemic:social) %>%
  mutate(weight = factor(weight, labels = c("epistemic", "social")),
         goal = factor(goal, levels = c("informative", "social", "both")))


# group_by(goal, param) %>%
#   summarize(MAP = estimate_mode(val),
#             cred_upper = hdi_upper(val),
#             cred_lower = hdi_lower(val))

# bda.state.tidy2 <- bda.utterance.results %>% 
#     select(contains("weight")) %>%
#     gather(key, val) %>% 
#     # mutate(key = gsub("value.", "", key)) %>% # for mht
#     separate(key, into = c("param", "goal", "state", "utterance")) %>%
#   mutate(weight = factor(param, labels = c("epistemic", "social")),
#          goal = factor(goal, levels = c("informative", "social", "both"))) %>%
#   ggplot(., aes(x=val, fill = weight))+
#   geom_density(alpha=0.5, adjust = 1.5, color ='black', size = 1.1)+
#     facet_grid(goal~., scales='free')+
#   theme_bw()+
#   ylab("posterior density")+
#   xlab("Inferred speaker goal weight")

ggplot(bda.state.tidy2, aes(x = val, fill = weight))+
  geom_histogram(binwidth=0.01, position = position_dodge())+
  facet_grid(goal~param) +
  # ylab("posterior density")+
  xlab("Inferred speaker goal weight")

# 
# ggplot(data=bda.state.tidy2, aes(x=val, fill = weight))+
#   geom_density(alpha=0.5, adjust = 0.5, color ='black', size = 0.5)+
#     facet_grid(goal~., scales='free')+
#   theme_bw()+
#   ylab("posterior density")+
#   xlab("Inferred speaker goal weight")
# 
  bda.state.tidy2 %>% 
  group_by(param, goal) %>%
  summarize(
    MAP = estimate_mode(val),
            ci_lower = hdi_lower(val),
            ci_upper = hdi_upper(val))
  
  # ggsave("~/Desktop/politeModelWeightsS1.pdf", width = 10)

```
      
      
### Predictives

#### scatter

```{r dataRenaming}
ms_data <- ms.all
levels(ms_data$goal)
levels(ms_data$goal) <- c("both", "informative", "social")
ms_data <- ms_data %>%
    mutate(MAP = as.numeric(as.character(mean)),
    ci_upper = as.numeric(as.character(ci_upper)),
    ci_lower = as.numeric(as.character(ci_lower))) %>%
  select(-mean) %>%
  mutate(goal = as.factor(goal),
         true_state = as.factor(true_state),
         positivity = as.factor(positivity),
         utterance = as.factor(utterance))

levels(ms_data$true_state) <- c(1:5)
levels(ms_data$positivity) <- c("neg", "no_neg")
```

```{r expModelTogether}

ms_model <- bda.utterance.tidy %>%     
  ungroup() %>%
  # mutate(goal = factor(goal, levels = c("informative", "social", "both"))) %>%
  mutate(true_state = state) %>%
  ungroup() %>%
  select(-state) %>%
  mutate(
    model_MAP = as.numeric(as.character(MAP)),
    model_ci_upper = as.numeric(as.character(ci_upper)),
    model_ci_lower = as.numeric(as.character(ci_lower))) %>%
  select(-MAP, -ci_upper, -ci_lower) %>%
  mutate(goal = as.factor(goal),
         true_state = as.factor(true_state),
         positivity = as.factor(positivity),
         utterance = as.factor(utterance))

ms_all <- left_join(ms_data, ms_model) %>%
  mutate(goal =factor(goal, levels=c("informative", "social", "both")),
         diff = (model_MAP - MAP)^2)
```

```{r expModelTogether_plot}
ggplot(ms_all, aes(x = model_MAP, y = MAP, col=goal)) +
  geom_point() +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper)) +
  geom_errorbarh(aes(xmin=model_ci_lower,xmax=model_ci_upper)) +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  xlab("Model posterior predictive") +
  ylab("Human proportion responses") +
  ylim(0,1) +
  xlim(0,1) +
  coord_fixed() +
  theme_bw()

# ggsave("~/Desktop/politeModelDataCorS2.pdf", width = 10)


# cor.test(ms_all$MAP, ms_all$model_MAP) # 0.8744156 ^2 = 0.7646
with(ms_all, cor(MAP, model_MAP))^2
```


```{r expModelTogether_plot2}
ggplot(ms_all, aes(x = model_MAP, y = MAP, col=goal)) +
  geom_point() +
#  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper)) +
 # geom_errorbarh(aes(xmin=model_ci_lower,xmax=model_ci_upper)) +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  xlab("Model posterior predictive") +
  ylab("Human proportion responses") +
  ylim(0,0.3) +
  xlim(0,0.3) +
  coord_fixed()

with(ms_all %>% filter(MAP < 0.3), cor(MAP, model_MAP))^2
```


### Negation vs Positive

```{r model_predictions_negNoneg}
bda.utterance.tidy4 <- bda.utterance.results %>%
    select(contains("negation")) %>%
    gather(key, val) %>%
     # mutate(key = gsub("value.", "", key)) %>% # for mht
    separate(key, into = c("param", "goal", "state", "positivity")) %>%
  mutate(val_noNeg = 1-val) %>%
  gather(posneg, val, val:val_noNeg) %>%
  select(-positivity) %>%
  mutate(positivity = factor(posneg, levels = c("val_noNeg", "val")),
         positivity = factor(positivity, labels = c("no negation", "negation"))) %>%
  mutate(goal = factor(goal, levels = c("informative", "social", "both"))) %>%
  select(-posneg) %>%
  group_by(goal, state, positivity) %>%
  # summarise(val = mean(val))
  summarize(MAP = estimate_mode(val),
            ci_lower = hdi_upper(val),
            ci_upper = hdi_lower(val))

ggplot(data=bda.utterance.tidy4, aes(x=state, y=MAP, col=positivity, group=positivity)) +
  # geom_bar(stat="identity", position=position_dodge()) +
  geom_line(stat="identity", position=position_dodge()) +
  facet_grid(goal~.) +
  xlab("true state") +
  ylab("proportion chosen") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper), position=position_dodge(width=.05)) +
  # ggtitle("negation (indirect remark) vs. no negation (white lie)")
  scale_colour_discrete(guide = guide_legend(title = "utterance type")) +
  theme_bw()

# ggsave("~/Desktop/politeModelNegNoneg.pdf", width = 5)

```
```{r modelData_negNoneg}
ms2 <- d %>%
  filter(!is.na(positivity), !is.na(utterance)) %>% # why is there NA?
  group_by(true_state, goal) %>%
  summarise(n.total=n())

ms3 <- d %>%
  filter(!is.na(positivity), !is.na(utterance)) %>% # why is there NA?
  group_by(true_state, goal, positivity) %>%
  summarize(n = n())

ms <- left_join(ms2, ms3) %>%
  group_by(true_state, goal, positivity) %>%
  summarize(mean = n / n.total,
            ci_lower = binom.bayes(n, n.total)$lower,
            ci_upper = binom.bayes(n, n.total)$upper) 
  
ms_fake <- expand.grid(true_state=levels(ms$true_state),
                             goal=levels(ms$goal),
                             positivity=levels(ms$positivity))

ms.all2 <- left_join(data.frame(ms_fake),data.frame(ms))

ms.all2[is.na(ms.all2)] <- 0
levels(ms.all2$true_state) <- c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts")
levels(ms.all2$goal) <- c("both", "informative", "social")
levels(ms.all2$positivity) <- c("negation", "no negation")

ms.all2 %<>%
  mutate(
         goal = fct_relevel(goal, "informative", "social", "both"))

ggplot(data=ms.all2, aes(x=true_state, y=mean, col=positivity, group=positivity)) +
  # geom_bar(stat="identity", position=position_dodge()) +
  geom_line(stat="identity", position=position_dodge()) +
  facet_grid(goal~.) +
  xlab("true state") +
  ylab("proportion chosen") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper), position=position_dodge(width=.05)) +
  # ggtitle("negation (indirect remark) vs. no negation (white lie)")
  scale_colour_discrete(guide = guide_legend(title = "utterance type")) +
  theme_bw()

ms_dataModNeg <- rbind(ms.all2 %>%
                         mutate(expt = "experiment") %>%
                         mutate(true_state = factor(true_state, labels = c(1:5))),
                       bda.utterance.tidy4 %>%
                         ungroup() %>%
                         mutate(true_state = state) %>%
                         # mutate(true_state = factor(state, labels = c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts"))) %>%
                         select(-state) %>%
                         mutate(mean = MAP) %>%
                         select(-MAP) %>%
                         mutate(expt = "model"))

ggplot(data=ms_dataModNeg, aes(x=true_state, y=mean, col=positivity, group=positivity)) +
  # geom_bar(stat="identity", position=position_dodge()) +
  geom_line(stat="identity", position=position_dodge()) +
  facet_grid(goal~expt) +
  xlab("true state") +
  ylab("proportion chosen") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper), position=position_dodge(width=.05)) +
  # ggtitle("negation (indirect remark) vs. no negation (white lie)")
  scale_colour_discrete(guide = guide_legend(title = "utterance type")) +
  theme_bw()

# ggsave("~/Desktop/politeModelDataNegS2.pdf", width = 6, height=8)

```

```{r model_predictions_eachUtt}
# bda.utterance.tidy2 <- bda.utterance.tidy %>%
#   filter(!is.na(val)) %>%
#   group_by(goal, state, positivity, utterance) %>%
#   distinct(val, .keep_all = TRUE)
# 
# bda.utterance.tidy3 <- bda.utterance.tidy %>%     
#   group_by(goal, state, positivity, utterance) %>%
#   # summarise(val = mean(val)) %>%
#   filter(!is.na(val)) %>%
#   summarize(
#     MAP = estimate_mode(val),
#             ci_lower = hdi_lower(val),
#             ci_upper = hdi_upper(val)) %>%
#   ungroup() %>%
#   mutate(goal = factor(goal, levels = c("informative", "social", "both"))) %>%
#   mutate(goal = factor(goal, labels = c("informative", "social", "Model pred for goal:both")))
#   
# bda.utterance.tidy3_org <- bda.utterance.tidy3 %>%
#   mutate(utterance = factor(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")),
#          state = factor(state, labels = c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts")),
#          positivity = factor(positivity, levels = c("no_neg", "neg")),
#          positivity = factor(positivity, labels = c("no negation", "negation")))

bda.utterance.tidy3_org <- bda.utterance.tidy %>%
  ungroup() %>%
  mutate(goal = factor(goal, levels = c("informative", "social", "both"))) %>%
  #mutate(goal = factor(goal, labels = c("informative", "social", "both"))) %>%
  mutate(utterance = factor(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")),
         state = factor(state, labels = c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts")),
         positivity = factor(positivity, levels = c( "neg", "no_neg"), labels = c("negation", "no negation")))

ggplot(data=bda.utterance.tidy3_org, aes(x=utterance, y=MAP, fill=positivity, color= positivity, group = positivity)) +
  #geom_bar(stat="identity", position=position_dodge()) +
  geom_line()+ 
  facet_grid(goal~state) +
  xlab("no neg (it was ~) vs neg (it wasn't ~) ") +
  ylab("proportion chosen") +
  ggtitle("Model predictions for speaker2 production (factor on both state and goals)") +
  geom_linerange(aes(ymin=ci_lower,ymax=ci_upper), position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  #theme_bw()+
  ylim(0,1)+
  scale_color_solarized()+
  ggthemes::theme_few()+
  #theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))

# ggsave("~/Desktop/politeModelS1.pdf", width = 10)

ggplot(data=ms.all, aes(x=utterance, y=mean, fill=positivity, color= positivity, group = positivity)) +
  #geom_bar(stat="identity", position=position_dodge()) +
  geom_line()+ 
  facet_grid(goal~true_state) +
  xlab("no neg (it was ~) vs neg (it wasn't ~) ") +
  ylab("proportion chosen") +
  ggtitle("Model predictions for speaker2 production (factor on both state and goals)") +
  geom_linerange(aes(ymin=ci_lower,ymax=ci_upper), position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  #theme_bw()+
  ylim(0,1)+
  scale_color_solarized()+
  ggthemes::theme_few()+
  #theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))

```

```{r expModelTogether-utt}
ms_dataMod <- rbind(ms_data %>% 
                      mutate(expt = "experiment") %>%
                      mutate(true_state = factor(true_state, labels = c("1 heart", "2 hearts", "3 hearts", "4 hearts", "5 hearts")),
                             positivity = factor(positivity, labels = c("negation", "no negation")),
                             goal = factor(goal, labels = c("informative", "social", "both"))), 
                    bda.utterance.tidy3_org %>% 
                      mutate(expt = "model") %>%
                      mutate(true_state = state) %>%
                      select(-state)) %>%
  mutate(expt = as.factor(expt)) %>%
  mutate(goal = factor(goal, levels = c("informative", "social", "both")))

ggplot(data=ms_dataMod, aes(x=utterance, y=MAP, fill=positivity, color= positivity, group = positivity)) +
  #geom_bar(stat="identity", position=position_dodge()) +
  geom_line()+ 
  facet_grid(goal+expt~true_state) +
  xlab("no neg (it was ~) vs neg (it wasn't ~) ") +
  ylab("proportion chosen") +
  # ggtitle("Model predictions for speaker2 production (factor on both state and goals)") +
  geom_linerange(aes(ymin=ci_lower,ymax=ci_upper), position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  #theme_bw()+
  ylim(0,1)+
  scale_color_solarized()+
  ggthemes::theme_few()+
  #theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))

 # ggsave("~/Desktop/politeDataModelS1.pdf", width = 10)

```



```{r check_chains_alphaOptimalityCost}
bda.utterance.results %>%
  filter(str_detect(Parameter, "speakerOptimality_NA_NA_NA|speakerOptimality2_NA_NA_NA|alpha_NA_NA_NA|cost_NA_NA_NA")) %>%
    separate(Parameter, into = c("param", "goal", "state", "utterance")) %>%
  ggplot(., aes(x = value, fill=as.factor(Chain)))+
  geom_histogram()+
  facet_grid(as.factor(Chain)~param, scales = "free")
```


```{r check_chains}
bda.utterance.tidy <- bda.utterance.results %>%
  filter(str_detect(Parameter, "posteriorPredictive")) %>%
  filter(!str_detect(Parameter, "negation")) %>%
    separate(Parameter, into = c("param", "goal", "state", "positivity", "utterance")) %>%
  mutate(utterance = factor(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")),
         positivity = factor(positivity, labels = c("neg", "no_neg")),
         positivity = factor(positivity, levels = c("no_neg", "neg"))) %>%
    group_by(Chain, goal, state, positivity, utterance) %>%
  filter(!is.na(value)) %>%
  # summarise(val = mean(val))
  summarize(MAP = estimate_mode(value),
            ci_lower = hdi_upper(value),
            ci_upper = hdi_lower(value))

ggplot(data=filter(bda.utterance.tidy, Chain == "1"), aes(x=positivity, y=MAP, fill=utterance)) +
  geom_bar(stat="identity", position=position_dodge()) +
  facet_grid(goal~state) +
  xlab("no neg (it was ~) vs neg (it wasn't ~) ") +
  ylab("proportion chosen") +
  ggtitle("Model predictions for speaker2 production (factor on both state and goals)") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper),position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  theme_bw()+
  ylim(0,1)

ggplot(data=filter(bda.utterance.tidy, Chain == "2"), aes(x=positivity, y=MAP, fill=utterance)) +
  geom_bar(stat="identity", position=position_dodge()) +
  facet_grid(goal~state) +
  xlab("no neg (it was ~) vs neg (it wasn't ~) ") +
  ylab("proportion chosen") +
  ggtitle("Model predictions for speaker2 production (factor on both state and goals)") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper),position="dodge") +
  geom_hline(yintercept=.1, lty=2) +
  theme_bw()+
  ylim(0,1)
```

```{r checkChains_cor}
ms_chain1 <- bda.utterance.tidy %>%
  filter(Chain == "1") %>%
  mutate(true_state = state) %>%
  ungroup() %>%
  select(-state) %>%
  mutate(
    chain1_MAP = as.numeric(as.character(MAP)),
    chain1_ci_upper = as.numeric(as.character(ci_upper)),
    chain1_ci_lower = as.numeric(as.character(ci_lower))) %>%
  select(-MAP, -ci_upper, -ci_lower, -Chain) %>%
  mutate(goal = as.factor(goal),
         true_state = as.factor(true_state),
         positivity = as.factor(positivity),
         utterance = as.factor(utterance))

ms_chain2 <- bda.utterance.tidy %>%
  filter(Chain == "2") %>%
  mutate(true_state = state) %>%
  ungroup() %>%
  select(-state) %>%
  mutate(
    chain2_MAP = as.numeric(as.character(MAP)),
    chain2_ci_upper = as.numeric(as.character(ci_upper)),
    chain2_ci_lower = as.numeric(as.character(ci_lower))) %>%
  select(-MAP, -ci_upper, -ci_lower, -Chain) %>%
  mutate(goal = as.factor(goal),
         true_state = as.factor(true_state),
         positivity = as.factor(positivity),
         utterance = as.factor(utterance))

ms_chains <- left_join(ms_chain2, ms_chain1) %>%
  mutate(goal =factor(goal, levels=c("informative", "social", "both")))

ggplot(ms_chains, aes(x = chain2_MAP, y = chain1_MAP, col=goal)) +
  geom_point() +
  geom_errorbar(aes(ymin=chain1_ci_lower,ymax=chain1_ci_upper)) +
  geom_errorbarh(aes(xmin=chain2_ci_lower,xmax=chain2_ci_upper)) +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  xlab("Model posterior predictive") +
  ylab("Human proportion responses") +
  ylim(0,0.95) +
  xlim(0,0.95) +
  coord_fixed()
```
```{r}
bda.utterance.results %>%
  filter(str_detect(Parameter, "posteriorPredictive")) %>%
  filter(!str_detect(Parameter, "negation")) %>%
    separate(Parameter, into = c("param", "goal", "state", "positivity", "utterance")) %>%
  mutate(utterance = factor(utterance, levels = c("terrible", "bad", "okay", "good", "amazing")),
         positivity = factor(positivity, labels = c("neg", "no_neg")),
         positivity = factor(positivity, levels = c("no_neg", "neg"))) %>%
    group_by(Chain, goal, state, positivity, utterance) %>%
  filter(!is.na(value)) %>%
    filter(Chain == "2") %>%
    ggplot(., aes( x = value, fill = goal, group = goal))+
  geom_histogram(position = position_dodge())+
  facet_grid(state ~ positivity + utterance, scales = 'free')
```

