---
title: "BDA of polite RSA"
author: "mht"
date: "November 11, 2015"
output: html_document
---

```{r helders}
library(coda)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   

fpath<- "~/Documents/research/polgrice/model/results/"
```


# Literal semantics data

```{r literal.sem.load}
fpath<- "~/Documents/research/polgrice/model/results/semantics/"
prefix<-"bda-semantics-thetaBernoulli-IncrMH"
prefix<- "bda-semantics-s1-nullAlt-IncrMH"
samples = 100000
burn =  samples /2
m<-read.csv(paste(fpath, prefix, samples/1000, "k-b", burn/1000,"k.csv", sep=''))
m.samples <- m[rep(row.names(m), m$Probability*samples), 1:3]
```


```{r semantics.bda}

s.tidy <- m.samples %>%
  group_by(Utterance,State) %>%
  summarise(MAP = estimate_mode(Theta),
            credHigh = HPDhi(Theta),
            credLow = HPDlo(Theta)) %>%
  ungroup() %>%
  mutate(Utterance = factor(Utterance, levels=c("terrible", "bad",
                                                "okay", "good",
                                                "amazing")))

ggplot(s.tidy, aes(x=State, y = MAP, color=Utterance))+
  geom_point()+
  geom_errorbar(aes(ymin=credLow, ymax=credHigh), width=0.1)+
  geom_line()+
  facet_grid(.~Utterance)

ggsave("~/Documents/research/polgrice/model/results/literalSemantics-bda-s1-negationAlt_incrmh100k-b50k.pdf")
```


# Model

```{r model.load}
prefix<- "bda-politeRSA-stochVal-sopt-goalBetaPriorsMH"
prefix<- "bda-politeRSA-stochWords-stochVal-sopt-goalBetaPriorsMH"
prefix<- "bda-politeRSA-stochWords-sopt-goalBetaPriorsMH"
prefix <- "bda-politeRSA-sopt-goalBetaPriorsMH"
prefix<- "bda-politeRSA-newNice-sopt-goalBetaPriorsMH"
prefix<-"bda-politeRSA-newNice-stochWords-sopt5-highKindPriorMH"
prefix<-"bda-politeRSA-newNice-stochWords-sopt5-uniformGoalsMH"
prefix<-"bda-politeRSA-jointUtilityProbScale-oneKindMean-goalPrior-sopt-MH"
prefix<-"bda-politeRSA-jointUtilityLogProbScale-sopt-MH"
prefix<-"bda-politeRSA-jointUtilityLogProbScale-sopt-MH"
prefix<-"bda-politeRSA-jointUtilityProbScale-OneSocialsopt-MH"
prefix<-"bda-politeRSA-jointUtilityProbScale-goalPriorFixed-sopt-MH"
prefix<-"bda-politeRSA-informativity-expectation-goalPrior-sopt-MH"
prefix<-"bda-politeRSA-widesopt-alphaMH"
# prefix<-"bda-politeRSA-jointUtilityMatchWeight-wMeanness-goalPrior-sopt-MH"
# prefix<-"bda-politeRSA-jointUtilityMatchWeight-wMeanness-goalPrior-sopt-MH"
prefix<-"bda-politeRSA-truesopt-alphaMH"
samples = 10000
burn =  samples /2
semsamp = 50000
semburn =semsamp/2
m.samples<- data.frame()
i<-"a"
for (i in c("a","b")){
m<-read.csv(paste(fpath, prefix, samples, "burn", burn,
                  "semantics-bern-",semsamp/1000,
                  "k-b",semburn/1000,"k",i,".csv", sep=''))
  m.samp <- m[rep(row.names(m), m$Probability*(samples-burn)), 1:5]
  m.samples<- bind_rows(m.samples, m.samp)
  print(i)
}
```


## Posterior over parameters

```{r posterior.parameters}

m.params <- m.samples %>%
  filter(Parameter=='parameter') %>%
  select(-Parameter, -Goal) %>%
  rename(Goal = State,
         Parameter = Utterance)

ggplot(m.params, aes(x=Value))+
  geom_histogram()+
  facet_grid(.~Goal+Parameter, scales='free')

m.params %>%
  group_by(Goal,Parameter) %>%
  summarize(MAP = estimate_mode(Value),
            credLow = HPDlo(Value),
            credHigh = HPDhi(Value))
# 
# ggsave(paste("~/Documents/research/polgrice/model/results/params",
#              prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
#                   "k-b",semburn/1000,"kx2.pdf", sep=''))
```

## Posterior predictive

```{r posterior.predictives}
  
lvls = c("terrible", "bad", "okay", "good", "amazing")
  
m.pp <- m.samples %>%
    filter(Parameter == 'predictive') %>%
    group_by(State, Utterance, Goal) %>%
    summarize(MAP = estimate_mode(Value),
              credHigh = HPDhi(Value),
              credLow = HPDlo(Value)) %>%
    ungroup() %>%
    mutate(#State = factor(State, levels = lvls, labels =c(1,2,3,4,5)),
           Utterance = factor(Utterance, levels = lvls))
  
#   
# write.csv(m.pp, 
#           file =paste("~/Documents/research/polgrice/model/results/goals",
#                       prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
#                   "k-b",semburn/1000,"k.csv", sep=""), row.names=F)

  
ggplot(m.pp, aes(x=State, y = MAP, color = Goal, group= Goal))+
    geom_point(size=2, position=position_dodge(0.3))+
    geom_line(position=position_dodge(0.3))+
    geom_errorbar(aes(ymin = credLow, ymax = credHigh),
                  position=position_dodge(0.3), width=0.2)+
    facet_grid(.~Utterance)

# ggsave(paste("~/Documents/research/polgrice/model/results/goals",
#              prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
#                   "k-b",semburn/1000,"k.pdf", sep='') )
```


### Compare to behavioral data

```{r}
d<-read.csv("~/Documents/research/polgrice/experiment/data_analysis/data/goal_unknown_rounded.csv")

d.tidy <- d %>%
  mutate(State = true_state,
         Utterance = utterance,
         Goal = goal) %>%
  mutate(State = factor(State, levels = lvls, labels =c(1,2,3,4,5))) %>%
  group_by(State, Utterance, Goal) %>%
  multi_boot_standard(column= "goal_prob")

md.tidy<- left_join(m.pp, d.tidy)

ggplot(md.tidy, aes(x=MAP, y = mean, color = Goal))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1, lty=2)+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper))+
  geom_errorbarh(aes(xmin=credLow, xmax = credHigh))+
  #theme_paper()+
  xlim(-0.05,1.05)+
  ylim(-0.05,1.05)+
  coord_fixed()

with(md.tidy, cor(MAP, mean))^2

```

```{r}
# To create model predictions for "mean" (for CUNY abstract)
m.pp.3<-bind_rows(m.pp %>%
            select(-credLow, -credHigh),
  
  m.pp %>% filter(Goal=='nice') %>%
  mutate(meanness = 1-MAP) %>%
  select(-credHigh, -credLow, -MAP, -Goal) %>%
  mutate(Goal = factor("mean")) %>%
  rename(MAP = meanness))


ggplot(m.pp.3, aes(x=Utterance, y = MAP, color = Goal, group= Goal))+
  geom_line(position=position_dodge(0.3))+
  geom_point(size=4)+
  #geom_errorbar(aes(ymin = credLow, ymax = credHigh), width=0.1)+
  facet_grid(.~State)

write.csv(m.pp.3, file = "~/Documents/research/polgrice/model/results/goals-sopt5-uniformGoals.csv", row.names=F)

```



# State given utterance and goal


```{r stateGivenUtt.model.load}
#prefix<- "bda-politeRSA-stateGivenUttGoal-3goalPrior-sopt-MH"
prefix<- "bda-politeRSA-stateGivenUttGoal-goalSpecPriorsAll3-sopt-MH"
#prefix<- "bda-politeRSA-stateGivenUttGoal-goalSpecPriors-sopt-MH"
prefix<-"bda-politeRSA-stateGivenUttGoal-informativity-expectation-goalPrior1Beta-sopt-MH"
prefix<-"bda-politeRSA-stateGivenUttGoal-goalPrior1Beta-sopt-alpha-phi-MH"
prefix<-"bda-politeRSA-stateGivenUttGoal-goalPrior1Beta-sopt-alpha-MH"
prefix<-"bda-politeRSA-stateGivenUttGoal-goalPrior1wide-Beta50d-sopt-alpha-IncrMH"
prefix<-"bda-politeRSA-stateGivenUttGoal-goal3Vals-Beta50d-sopt-alpha-IncrMH"
samples = 100000
burn =  samples /2
semsamp = 50000
semburn =semsamp/2
# m<-read.csv(paste(fpath, prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
# #                  "semantics-S1-negAlt-",semsamp/1000,
# #                  "semantics-S1-nullAlt-",semsamp/1000,
#                   "k-b",semburn/1000,"kb.csv", sep=''))
# m.samples <- m[rep(row.names(m), m$Probability*samples), 1:5]
# 
i<-"c"
m.samples<- data.frame()
for (i in c("a", "b")){
m<-read.csv(paste(fpath, prefix, samples, "burn", burn,
                  "semantics-bern-",semsamp/1000,
                  "k-b",semburn/1000,"k",i,".csv", sep=''))
  m.samp <- m[rep(row.names(m), m$Probability*(samples-burn)), 1:5]
  m.samples<- bind_rows(m.samples, m.samp)
  print(i)
}
```


## Posterior over parameters

```{r posterior.parameters}

m.params <- m.samples %>%
  filter(Parameter=='parameter') %>%
  #select(-Parameter, -Goal) %>%
  select(-Parameter) %>%
  rename(Cndition = State,
           #Goal = State,
         Goal = Utterance,
         Parameter = Goal) %>%
  mutate(Cndition = factor(Cndition, levels=c("honest", "nice", "mean")),
         Goal = factor(Goal, levels=c("honesty", "kindness", "meanness"),
                       labels=c("honesty", "niceness", "meanness")))

ggplot(m.params, aes(x=Value, fill = Goal))+
  geom_histogram(position=position_dodge())+
  facet_grid(Cndition~Parameter, scales='free')

View(m.params %>%
  group_by(Cndition, Goal,Parameter) %>%
  summarize(MAP = estimate_mode(Value),
            credLow = HPDlo(Value),
            credHigh = HPDhi(Value)))

ggsave(paste("~/Documents/research/polgrice/model/results/params",
             prefix, samples, "burn", burn,
                  "semantics-bern-",semsamp/1000,
                  "k-b",semburn/1000,"k.pdf", sep=''), width =12, height = 8)



ggplot(m.params %>% filter(Parameter=="weight"), aes(x=Value, fill = Goal))+
  geom_density(alpha=0.5, adjust = 3.5, color ='black', size = 1.1)+
    facet_grid(Cndition~., scales='free')+
  theme_bw()+
  ylab("posterior density")+
  xlab("Inferred speaker goal weight")

ggsave("~/Documents/research/polgrice/conference/cogsci/abstract/figures/goal-posterior.pdf", height =5, width = 5)

```



## Posterior predictive

```{r posterior.predictives}

lvls = c("terrible", "bad", "okay", "good", "amazing")

m.pp <- m.samples %>%
  filter(Parameter == 'predictive' & Goal == 'expectation') %>%
  select(-Goal) %>%
  rename(Goal = State) %>%
  mutate(Value = Value/20) %>%
  group_by(Utterance, Goal) %>%
  summarize(MAP = estimate_mode(Value),
            credHigh = HPDhi(Value),
            credLow = HPDlo(Value)) %>%
  ungroup() %>%
  mutate(#State = factor(State, levels = lvls, labels =c(1,2,3,4,5)),
         Utterance = factor(Utterance, levels = lvls),
         Goal = factor(Goal, levels =c("honest", "nice", "mean")),
         src = "model") %>%
  rename(mean = MAP,
         ci_lower = credLow,
         ci_upper = credHigh)

# 
# write.csv(m.pp, 
#           file =paste("~/Documents/research/polgrice/model/results/states",
#                       prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
#                   "k-b",semburn/1000,"k.csv", sep=""), row.names=F)

fig1<-ggplot(m.pp, aes(x=Utterance, y = mean, color = Goal, group= Goal))+
  geom_point(size=2, position=position_dodge(0.3))+
  geom_line(position=position_dodge(0.3))+
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position=position_dodge(0.3), width=0.2)+
  ylim(1,5)+
  ylab("Model posterior predictive")
  theme_bw()

fig1
# 
# ggsave(paste("~/Documents/research/polgrice/model/results/goals",
#              prefix, samples, "burn", burn,
#                   "semantics-bern-",semsamp/1000,
#                   "k-b",semburn/1000,"k.pdf", sep=''), width = 8, height = 5 )
```

Figure of Expectation (data and model) 

```{r}
ds.tidy <- d.s %>%
  mutate(State = state/20,
         Utterance = utterance,
         Goal = goal) %>%
  #mutate(State = factor(State, levels = lvls, labels =c(1,2,3,4,5))) %>%
  group_by(Utterance, Goal) %>%
  multi_boot_standard(column= "State") %>%
  mutate(src = 'experiment')


md.bars <- bind_rows(m.pp, ds.tidy)
fig1<-ggplot(md.bars, aes(x=Utterance, y = mean, color = Goal, group= Goal))+
  geom_point(size=3, position=position_dodge(0.3))+
  geom_line(position=position_dodge(0.3))+
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position=position_dodge(0.3), width=0.2)+
  ylim(1,5)+
  ylab("Inferred true state")+
  facet_grid(~src)+
  theme_bw()+
  theme(legend.title = element_text(hjust=0),
        legend.position="bottom",
        legend.direction="horizontal")

ggsave("~/Documents/research/polgrice/conference/cogsci/abstract/figures/state-inference.pdf", height =4, width = 6)


 md.tidy<- left_join(m.pp %>% select(-src), ds.tidy)

ggplot(md.tidy, aes(x=MAP, y = mean, color = Goal))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1, lty=2)+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper))+
  geom_errorbarh(aes(xmin=credLow, xmax = credHigh))+
  theme_paper()+
  coord_fixed()+
  xlim(0,100)+
  ylim(0,100)

with(md.tidy%>% filter(Goal!='mean')
     , cor(mean, MAP))^2
```


### Full distribution


### Model - data fit

Full distribution
```{r}



d.s<-read.csv("~/Documents/research/polgrice/model/data/state.csv")

ds.dist <- d.s %>%
  rename(State = state,
         Utterance = utterance,
         Goal = goal) %>%
    mutate(State = State/20) %>%
  #mutate(State = factor(State, levels = lvls, labels =c(1,2,3,4,5))) %>%
  group_by(Utterance, Goal, State) %>%
  summarize(n = length(subid)) %>%
  ungroup() %>%
  group_by(Utterance, Goal) %>%
  mutate(norm_n = n / sum(n))


m.pp <- m.samples %>%
  filter(Parameter == 'predictive' & !(Goal == 'expectation')) %>%
  rename(State1 = Goal,
         Goal = State) %>%
  mutate(State1 = to.n(State1)/20) %>%
  group_by(Utterance,Goal, State1) %>%
  summarize(MAP = estimate_mode(Value),
            credHigh = HPDhi(Value),
            credLow = HPDlo(Value)) %>%
  ungroup() %>%
  mutate(#State = factor(State, levels = lvls, labels =c(1,2,3,4,5)),
         Utterance = factor(Utterance, levels = lvls),
         Goal = factor(Goal, levels =c("honest", "nice", "mean"))) %>%
  rename(State = State1)

md.dist<-left_join(m.pp, ds.dist)

md.dist[is.na(md.dist$n),"norm_n"]<-0

md.dist$State = factor(md.dist$State, levels=c(1,2,3,4,5))
md.dist$Goal = factor(md.dist$Goal, levels=c("honest", "nice", "mean"))

ggplot(md.dist, aes(x = MAP, y = norm_n, 
                    xmin = credLow, xmax = credHigh, 
                    color = Goal))+
  geom_point(size = 2)+
  #scale_alpha_manual(values = c(0.5, 0.8, 0.8, 0.8, 1))+
  geom_errorbarh()+
  geom_abline(intercept=0, slope=1, lty = 3)+
  coord_fixed()+
  xlim(-0.05,0.9)+
  ylim(-0.05,0.9)+
  xlab("Model posterior predictive")+
  ylab("Human proportion responses")+
  theme_bw()+
  guides(color=F)
#ggsave("~/Documents/research/polgrice/conference/cogsci/abstract/figures/state-inference-scatter.pdf", height =4, width = 4)


with(md.dist %<% filter(Goal!="mean"),
     cor(MAP, norm_n))^2
```



### Predictive over goal weight priors
 (the cheap way)
Reconstruct goal weight prior by forward sampling

```{r posterior.predictive.forwardSample}



m.prior.params <- m.params %>% filter(!(is.na(Goal))) %>%
  filter((Cndition=='honest' & Goal=='honesty') |
           (Cndition=='nice' & Goal == "kindness") |
           (Cndition=='mean' & Goal == 'meanness'))

shape1 = function(gamma,delta){return (gamma * delta)}
shape2= function(gamma,delta){return ((1-gamma) * delta)}

d.samples<-data.frame()

for (i in 1:1000){

  d.samp<-m.prior.params %>%
    group_by(Cndition, Goal, Parameter) %>%
    sample_n(1) %>%
    ungroup() %>%
    spread(Parameter, Value) %>%
    mutate(alpha = shape1(gamma,delta),
           beta = shape2(gamma, delta))
  
    for (k in 1:length(d.samp$Goal)){
    d.iter<-data.frame(
               Cndition = d.samp[k,1],
               Goal = d.samp[k,2],
               Weight = rbeta(10, d.samp[[k, "alpha"]], d.samp[[k, "beta"]])
    )
    d.samples<-bind_rows(d.samples, d.iter)
  }
  
#   %>%
#     group_by(Cndition, Goal)
#   
#   
#   %>%
#     mutate(goalweight = rbeta(10, alpha, beta)) %>%
#     mutate(hasDoneA = rbinom(1,1,prob_hasF)) %>%
#     select(-gamma, -delta, -Measure)
# 
#   
#     d.freq.samp<-m2.params %>%
#     group_by(Item, Gender, Parameter) %>%
#     sample_n(1) %>%
#     ungroup() %>%
#       spread(Parameter, Value) %>%
#       select(-Measure)
#     
#     d.samp <- left_join(d.samp.exis, d.freq.samp)
#   

  
  if ((i %% 10)==0) { print(i) }
}

ggplot(d.samples, aes(x=Weight, fill=Goal))+
  geom_density(adjust=0.9,alpha = 0.6, color='black',size = 1.1)+
  #facet_wrap(~Goal)+
  guides(color=F)+
  scale_x_continuous(breaks=c(0,0.5,1))

ggsave("~/Documents/research/polgrice/conference/cogsci/abstract/figures/goal-priors-bda.pdf")

```
