---
title: "Negation politeness"
author: "Erica Yoon"
date: "Sep 30 2016"
output: html_document
---

# Literal semantics

load data.

```{r warning=FALSE, message=FALSE}
rm(list = ls())
library(jsonlite)
library(ggplot2)
library(tidyr)
library(dplyr)
library(binom)
library(bootstrap)
library(langcog)

raw.data.path <- "/Users/ericang/Documents/Research/polgrice_GIT/experiment/exp_versions/negimp_prior_v1/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```

Filter out participants and clean up.

```{r}
d <- all.data %>%
  select(subid, answers.data.order, answers.data.domain, answers.data.state, answers.data.utterance, answers.data.judgment) %>%
  mutate(subid = as.factor(subid)) %>%
  mutate(trial = as.numeric(answers.data.order)) %>%
  mutate(item = as.factor(answers.data.domain)) %>%
  mutate(state = as.factor(substr(answers.data.state, 6, 6))) %>%
  mutate(utterance = as.factor(answers.data.utterance)) %>%
  mutate(judgment = as.numeric(as.character(answers.data.judgment))) %>%
  mutate(positivity = factor(as.numeric(grepl("yes", utterance)), 
                        levels = c(0, 1), 
                        labels = c("negative","positive"))) %>%
  mutate(utterance = substring(utterance, 5)) %>%
  select(subid, trial, item, state, positivity, utterance, judgment)

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
d$positivity <- relevel(d$positivity, ref="positive")
levels(d$positivity) <- c("it was ___", "it wasn't ___")
```

judgment ~ state + utterance + positivity (was vs. wasn't)

```{r echo = FALSE, fig.width = 12, fig.height = 4}
ms <- d %>%
  group_by(positivity, state, utterance, subid) %>%
  summarize(
            judgment = mean(judgment, na.rm=TRUE)
          ) %>%
  group_by(positivity, state, utterance) %>%
  multi_boot_standard(column = "judgment") %>%
  mutate(judgment = mean)

qplot(state, judgment, 
      colour = positivity,
      data=ms) + 
  geom_line(aes(group=positivity)) +
  facet_grid(.~utterance) +
  xlab("state (1=worst)") +
  ylab("proportion of acceptances") +
  ggtitle("Literal semantics") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  scale_color_discrete(guide_legend(title=""))
```

# (S & G | U)

Given utterance, infer state, 
and then given utterance and state, infer goals

```{r warning=FALSE, message=FALSE, echo=FALSE}
rm(list = ls())
library(jsonlite)
library(ggplot2)
library(tidyr)
library(dplyr)
library(binom)
library(bootstrap)
library(langcog)
source("/Users/ericang/Documents/Research/polgrice_GIT/experiment/data_analysis/helper/useful.R")

raw.data.path <- "/Users/ericang/Documents/Research/polgrice_GIT/experiment/exp_versions/7_negimp_v1/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```

clean up.

```{r}
rearr <- all.data %>%
  select(subid, num_range("answers.data.goal", 0:1)) %>%
  distinct(subid) %>%
  gather(goal_order, goal, num_range("answers.data.goal", 0:1)) %>%
  mutate(goal_order = substr(goal_order, 18, 18))

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.utterance, answers.data.judgment, answers.data.state, answers.data.threat, num_range("answers.data.goalProb", 0:1)) %>%
  gather(goal_order, goal_prob, num_range("answers.data.goalProb", 0:1)) %>%
  mutate(goal_order = substr(goal_order, 22, 22)) %>%
  left_join(rearr)


d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(actual_state = substr(answers.data.state, 6, 6)) %>%
  mutate(threat = answers.data.threat) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  mutate(predicted_state = answers.data.judgment) %>%
  select(subid, trial, knowledge, utterance, threat, goal, predicted_state, actual_state, goal_prob)

d <- d %>%
  mutate(subid = as.factor(subid),
         trial = as.numeric(trial),
         predicted_state = as.numeric(predicted_state)/20,
         actual_state = as.factor(actual_state),
         knowledge = as.factor(knowledge),
         threat = as.factor(threat),
         utterance = as.factor(utterance),
         goal = as.factor(goal),
         goal_prob = as.numeric(goal_prob)) %>%
  mutate(positivity = factor(as.numeric(grepl("yes", utterance)), 
                        levels = c(0, 1), 
                        labels = c("negative","positive"))) %>%
  mutate(utterance = substring(utterance, 5))

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
d$goal <- ordered(d$goal, levels = c("informative", "polite"))

d_neg <- d
```

Number of participants

```{r}
subj <- d_neg %>%
  group_by(threat) %>%
  summarise(count = length(unique(subid))) 

knitr::kable(subj)
```


also load data from original S and G expts (given the other parameter - G or S).

```{r warning=FALSE, message=FALSE}
### L2_S

raw.data.path <- "/Users/ericang/Documents/Research/polgrice_GIT/experiment/exp_versions/6_L2_S/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  jso1$site <- "US"
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.goal, answers.data.utterance, answers.data.judgment)

d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  mutate(goal = answers.data.goal) %>%
  mutate(state = answers.data.judgment) %>%
  select(subid, trial, knowledge, utterance, goal, state)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$state <- as.factor(d$state) 
d$knowledge <- as.factor(d$knowledge)
d$utterance <- as.factor(d$utterance)
d$goal <- as.factor(d$goal)
d$state <- as.numeric(d$state)

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
d$goal <- ordered(d$goal, levels = c("honest", "nice", "mean"))

d_S <- d

### L2_G

raw.data.path <- "/Users/ericang/Documents/Research/polgrice_GIT/experiment/exp_versions/5_L2_G/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}

rearr <- all.data %>%
  select(subid, num_range("answers.data.goal", 0:2)) %>%
  distinct(subid) %>%
  gather(goal_order, goal, num_range("answers.data.goal", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 18, 18))

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.state, answers.data.utterance, num_range("answers.data.goalProb", 0:2)) %>%
  gather(goal_order, goal_prob, num_range("answers.data.goalProb", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 22, 22))

detach(package:dplyr)
library(plyr)
d <- join(d, rearr)

detach(package:plyr)
library(dplyr)
d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(true_state = substr(answers.data.state, 6, 6)) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  select(subid, trial, knowledge, true_state, utterance, goal, goal_prob)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$true_state <- as.factor(d$true_state) 
d$knowledge <- as.factor(d$knowledge)
d$utterance <- as.factor(d$utterance)
d$goal <- as.factor(d$goal)
d$goal_prob <- as.numeric(d$goal_prob)

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))

d_G <- d
```

looking at negation politeness: just no identity condition.

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 5, fig.height = 4}
d <- d_neg

# predicted_state ~ utterance + positivity
ms <- d %>%
  group_by(threat, utterance, positivity, subid) %>%
  summarise(predicted_state = mean(predicted_state)) %>%
  group_by(threat, utterance, positivity) %>%
  multi_boot_standard(column = "predicted_state") %>%
  mutate(predicted_state = mean)
ms$positivity <- relevel(ms$positivity, ref="positive")
levels(ms$positivity) <- c("it was ___", "it wasn't ___")
levels(ms$threat) <- c("no identity", "face threat", "no face threat")

ggplot(ungroup(ms %>% filter(threat == "no identity")), aes(x=utterance, y = predicted_state, 
      colour = positivity)) +
  geom_line(aes(group = positivity)) +
  geom_point(aes(group = positivity)) +
  xlab("utterance") +
  ylab("inferred state") +
  ggtitle("Inferred state given utterance and goal") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  scale_color_discrete(guides(title=""))

```

no identity vs. face threat.

```{r}
ggplot(ungroup(ms %>% filter(threat != "no face threat")), aes(x=utterance, y = predicted_state, 
      colour = threat,
      linetype = positivity)) +
  geom_line(aes(group = interaction(positivity, threat))) +
  geom_point(aes(group = threat)) +
  xlab("utterance") +
  ylab("inferred state") +
  ggtitle("Inferred state given utterance and goal") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  scale_linetype_discrete(guides(title="")) +
  scale_color_discrete(guides(title=""))
```

compare to original S expt.

```{r}
d_comp <- d %>%
  mutate(state = predicted_state) %>%
  filter(positivity == "positive", 
         threat == "no_identity") %>% # just no identity
  select(subid, trial, goal, state, utterance) %>%
  mutate(expt = "new:no_threat")

d_S_honest <- d_S %>%
  filter(goal == "honest") %>%
  select(subid, trial, goal, state, utterance) %>%
  mutate(expt = "original:threat_honest")

d_S_nice <- d_S %>%
  filter(goal == "nice") %>%
  select(subid, trial, goal, state, utterance) %>%
  mutate(expt = "original:threat_nice")
  
  
d_S_comp <- rbind(d_comp, d_S_honest, d_S_nice)

ms <- d_S_comp %>%
  group_by(expt, utterance, subid) %>%
  summarise(state = mean(state)) %>%
  group_by(expt, utterance) %>%
  multi_boot_standard(column = "state") %>%
  mutate(state = mean)

ggplot(ms, aes(x=utterance, y = state, 
      colour = expt)) +
  geom_line(aes(group = expt)) +
  geom_point(aes(group = expt)) +
  xlab("utterance") +
  ylab("inferred state") +
  ggtitle("S|G&U (with face threat) vs. S|U (no identity)") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  scale_color_discrete(guides(title=""))
```

inferred meaning of positive utterance (e.g. it was terrible) in no identity condition looks like honest condition in original expt.


goal inferences.

```{r warning=FALSE, message=FALSE, fig.width = 10, fig.height = 3}
ms <- d %>%
  filter(threat == "no_identity") %>%
  group_by(goal, utterance, positivity, actual_state, subid) %>%
  summarise(goal_prob = mean(goal_prob)) %>%
  group_by(goal, utterance, positivity, actual_state) %>%
  multi_boot_standard(column = "goal_prob") %>%
  mutate(goal_prob = mean)

ms$positivity <- relevel(ms$positivity, ref="positive")
levels(ms$positivity) <- c("it was ___", "it wasn't ___")

qplot(actual_state, goal_prob, 
      colour = goal,
      data=ms) + 
  geom_line(aes(group=goal)) +
  facet_grid(positivity~utterance) +
  xlab("actual_state") +
  ylab("inferred goal") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  ggtitle("Goal inferences")
```

Compare with original G experiment.

```{r, fig.width = 10, fig.height = 5}
d_G_comp0 <- d_G %>%
  select(subid, trial, true_state, utterance, goal, goal_prob) %>%
  mutate(expt = "new:threat")

d_G_comp1 <- d %>%
  filter(positivity == "positive", 
         threat == "no_identity") %>% # just no identity
  mutate(true_state = actual_state) %>%
  select(subid, trial, true_state, utterance, goal, goal_prob) %>%
  mutate(expt = "original:no_threat",
         goal = factor(goal, labels = c("honest", "nice")))
  
d_G_comp <- rbind(d_G_comp0, d_G_comp1)

ms <- d_G_comp %>%
  group_by(expt, goal, utterance, true_state, subid) %>%
  summarise(goal_prob = mean(goal_prob)) %>%
  group_by(expt, goal, utterance, true_state) %>%
  multi_boot_standard(column = "goal_prob") %>%
  mutate(goal_prob = mean)

qplot(true_state, goal_prob, 
      colour = expt,
      data=filter(ms, goal != "mean")) + 
  geom_line(aes(group=expt)) +
  facet_grid(goal~utterance) +
  xlab("actual_state") +
  ylab("inferred goal") +
  geom_errorbar(aes(ymin=ci_lower,ymax=ci_upper,width=.1)) +
  ggtitle("Compare new (no threat) vs. original (threat)")
```
