---
title: "polgrice-goal v3 ana"
author: "Erica Yoon"
date: "October 24, 2015"
output: html_document
---

```{r warning=FALSE, message=FALSE, echo=FALSE}
rm(list = ls())
library(jsonlite)
library(ggplot2)
library(tidyr)
library(dplyr)
library(binom)
library(bootstrap)
source("/Users/ericang/Documents/Research/polite_probmods/experiment/data_analysis/helper/useful.R")

raw.data.path <- "/Users/ericang/Documents/Research/polite_probmods/polite_adj/experiment/exp_versions/3_goals_v3/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, }
rearr <- all.data %>%
  select(subid, num_range("answers.data.goal", 0:2)) %>%
  distinct(subid) %>%
  gather(goal_order, goal, num_range("answers.data.goal", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 18, 18))

rearr2 <- all.data %>%
  select(subid, num_range("answers.data.prediction", 0:1)) %>%
  distinct(subid) %>%
  gather(pred_order, prediction, num_range("answers.data.prediction", 0:1)) %>%
  mutate(pred_order = substr(pred_order, 24, 24))

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.state, answers.data.utterance, num_range("answers.data.predictedProb", 0:1), num_range("answers.data.inferredProb", 0:2)) %>%
  gather(pred_order, pred_prob, num_range("answers.data.predictedProb", 0:1)) %>%
  mutate(pred_order = substr(pred_order, 27, 27)) %>%
  gather(goal_order, goal_prob, num_range("answers.data.inferredProb", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 26, 26))


detach(package:dplyr)
library(plyr)
d <- join(d, rearr)
d <- join(d, rearr2)

detach(package:plyr)
library(dplyr)
d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(true_state = answers.data.state) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  select(subid, trial, knowledge, true_state, utterance, goal, goal_prob, prediction, pred_prob)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$true_state <- as.factor(d$true_state) 
d$knowledge <- as.factor(d$knowledge)
d$utterance <- as.factor(d$utterance)
d$goal <- as.factor(d$goal)
d$prediction <- as.factor(d$prediction)
d$goal_prob <- as.numeric(d$goal_prob)
d$pred_prob <- as.numeric(d$pred_prob)

d$true_state <- ordered(d$true_state, levels = c("terrible", "bad", "okay", "good", "amazing"))
d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
```

```{r echo=FALSE, fig.width = 10, fig.height = 4}
## for bootstrapping 95% confidence intervals
theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}
```

# goal_prob ~ true_state + utterance + goal + knowledge

```{r echo=FALSE, fig.width = 10, fig.height = 4}
# goal_prob ~ true_state + utterance + goal + knowledge
mss <- d %>%
  group_by(true_state, utterance, goal, knowledge, subid) %>%
  summarize(
            goal_prob = mean(goal_prob, na.rm=TRUE)
          )
ms <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, mean)
ms$cih <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.high)$goal_prob
ms$cil <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.low)$goal_prob
```

```{r echo = FALSE, fig.width = 12, fig.height = 4}
ms$goal <- ordered(ms$goal, levels = c("honest", "nice", "mean"))
levels(ms$true_state) <- c("state: terrible", "state: bad", "state: okay", "state: good", "state: amazing")

# unknown
qplot(utterance, goal_prob, 
      colour = goal,
      data=subset(ms, knowledge == "unknown")) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~true_state) +
  xlab("utterance") +
  ylab("likelihood for goal") +
  ggtitle("What was the speaker's goal? (with state-utt match)") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

# unknown
levels(ms$true_state) <- c("terrible", "bad", "okay", "good", "amazing")
qplot(true_state, goal_prob, 
      colour = utterance,
      data=subset(ms, knowledge == "unknown" & goal=="honest")) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~utterance) +
  xlab("state") +
  ylab("likelihood for goal") +
  ggtitle("honesty goal given state and utterance") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

# histogram
qplot(goal_prob, data=subset(ms, knowledge == "unknown" & true_state == "terrible")) + 
  geom_histogram(binwidth = 0.25) +
  facet_wrap(~utterance, ncol=5)

# known
qplot(utterance, goal_prob, 
      colour = goal,
      data=subset(ms, knowledge == "known")) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~true_state) +
  xlab("utterance") +
  ylab("likelihood for goal") +
  ggtitle("What was the speaker's goal? (listener knowledgeable)") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))
```

# pred_prob ~ true_state + utterance + prediction + knowledge

```{r echo = FALSE, fig.width = 10, fig.height = 4}
# prob_ask ~ true_state + utterance
mss <- d %>%
  group_by(true_state, utterance, prediction, knowledge, subid) %>%
  summarize(
            pred_prob = mean(pred_prob, na.rm=TRUE)
          )
ms <- aggregate(pred_prob ~ true_state + utterance + prediction + knowledge, mss, mean)
ms$cih <- aggregate(pred_prob ~ true_state + utterance + prediction + knowledge, mss, ci.high)$pred_prob
ms$cil <- aggregate(pred_prob ~ true_state + utterance + prediction + knowledge, mss, ci.low)$pred_prob

# plot
qplot(utterance, pred_prob, 
      data=subset(ms, knowledge == "unknown"),
      colour = prediction) + 
  geom_line(aes(group=prediction)) +
  facet_grid(.~true_state) +
  xlab("utterance") +
  ylab("likelihood") +
  scale_y_continuous(limits = c(0, 1)) +
  ggtitle("How likely is it for you to ask this person for feedback? / like this person? (listener not knowledgeable)") +
  geom_errorbar(aes(ymin=pred_prob-cil,ymax=pred_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

qplot(utterance, pred_prob, 
      data=subset(ms, knowledge == "known"),
      colour = prediction) + 
  geom_line(aes(group=prediction)) +
  facet_grid(.~true_state) +
  xlab("utterance") +
  ylab("likelihood") +
  scale_y_continuous(limits = c(0, 1)) +
  ggtitle("How likely is it for you to ask this person for feedback? / like this person? (listener knowledgeable)") +
  geom_errorbar(aes(ymin=pred_prob-cil,ymax=pred_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))


```

# correlations

```{r}
d1 <- d
d1$true_state <- as.numeric(d1$true_state)
d1$utterance <- as.numeric(d1$utterance)
d1 <- d1 %>% 
  mutate(distance = true_state-utterance)

summary(lmer(goal_prob ~ abs(distance) + (1 | subid), data=subset(d1,knowledge == "unknown" & goal == "honest")))
summary(lmer(goal_prob ~ true_state * utterance + (1 | subid), data=subset(d1,knowledge == "unknown" & goal == "nice")))
summary(lmer(goal_prob ~ true_state * utterance + (1 | subid), data=subset(d1,knowledge == "unknown" & goal == "mean")))
    

d1 <- d %>% filter(prediction == "like")
cor(d1[d1$goal=="nice",]$pred_prob, d1[d1$goal=="nice",]$goal_prob)
cor(d1[d1$goal=="mean",]$pred_prob, d1[d1$goal=="mean",]$goal_prob)
cor(d1[d1$goal=="honest",]$pred_prob, d1[d1$goal=="honest",]$goal_prob)

d1 <- d %>% filter(prediction == "ask")
cor(d1[d1$goal=="nice",]$pred_prob, d1[d1$goal=="nice",]$goal_prob)
cor(d1[d1$goal=="mean",]$pred_prob, d1[d1$goal=="mean",]$goal_prob)
cor(d1[d1$goal=="honest",]$pred_prob, d1[d1$goal=="honest",]$goal_prob)

```


```{r}
# model
model <- read.csv("Documents/Research/polite_probmods/polite_adj/model/results/goals-sopt5-uniformGoals.csv")
model$State <- factor(model$State, levels = c("terrible", "bad", "okay", "good", "amazing"))
levels(model$State) <- c("state: terrible", "state: bad", "state: okay", "state: good", "state: amazing")
model$Utterance <- factor(model$Utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
model$Goal <- factor(model$Goal, levels = c("honest", "nice", "mean"))

model <- model %>%
  mutate(true_state = State) %>%
  mutate(utterance = Utterance) %>%
  mutate(goal = Goal) %>%
  mutate(goal_prob = MAP) 
model$State <- NULL
model$Utterance <- NULL
model$Goal <- NULL
model$MAP <- NULL

model$cih <- NA
model$cil <- NA
model$knowledge <- NA
model$cat <- "model"

qplot(utterance, goal_prob, 
      colour = goal,
      data=model) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~true_state) +
  xlab("utterance") +
  ylab("likelihood for goal") +
  ggtitle("Model predictions") +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

# goal_prob ~ true_state + utterance + goal + knowledge
mss <- d %>%
  group_by(true_state, utterance, goal, knowledge, subid) %>%
  summarize(
            goal_prob = mean(goal_prob, na.rm=TRUE)
          )
ms <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, mean)
ms$cih <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.high)$goal_prob
ms$cil <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.low)$goal_prob
levels(ms$true_state) <- c("state: terrible", "state: bad", "state: okay", "state: good", "state: amazing")

ms$cat <- "experiment"


modeld <- rbind(subset(ms, knowledge == "unknown"), model)
modeld$cat <- as.factor(modeld$cat)
modeld$cat <- relevel(modeld$cat, "model")
qplot(utterance, goal_prob, 
      colour = goal,
      data=modeld) + 
  geom_line(aes(group=goal), position=position_dodge()) +
  facet_grid(cat~true_state) +
  xlab("utterance") +
  ylab("likelihood for goal") +
  ggtitle("What was the speaker's goal?") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))
