---
title: "Summary of expts so far"
author: "Erica Yoon"
date: "Dec 3, 2015"
output: html_document
---

# Three expts
(1) infer goals given utterance and word-scale state
(2) infer goals given utterance and heart-scale state 
(3) accept word (feeling) given heart-scale state

```{r warning=FALSE, message=FALSE, echo=FALSE}
rm(list = ls())
library(jsonlite)
library(ggplot2)
library(tidyr)
library(dplyr)
library(binom)
library(bootstrap)
source("/Users/ericang/Documents/Research/polgrice_GIT/experiment/data_analysis/helper/useful.R")

raw.data.path <- "/Users/ericang/Documents/Research/polgrice_GIT/experiment/exp_versions/5_L2_G/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
rearr <- all.data %>%
  select(subid, num_range("answers.data.goal", 0:2)) %>%
  distinct(subid) %>%
  gather(goal_order, goal, num_range("answers.data.goal", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 18, 18))

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.state, answers.data.utterance, num_range("answers.data.goalProb", 0:2)) %>%
  gather(goal_order, goal_prob, num_range("answers.data.goalProb", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 22, 22))

detach(package:dplyr)
library(plyr)
d <- join(d, rearr)

detach(package:plyr)
library(dplyr)
d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(true_state = substr(answers.data.state, 6, 6)) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  select(subid, trial, knowledge, true_state, utterance, goal, goal_prob)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$true_state <- as.factor(d$true_state) 
d$knowledge <- as.factor(d$knowledge)
d$utterance <- as.factor(d$utterance)
d$goal <- as.factor(d$goal)
d$goal_prob <- as.numeric(d$goal_prob)

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
## for bootstrapping 95% confidence intervals
theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}

# goal_prob ~ true_state + utterance + goal
mss <- d %>%
  group_by(true_state, utterance, goal, subid) %>%
  summarize(
            goal_prob = mean(goal_prob, na.rm=TRUE)
          )
ms <- aggregate(goal_prob ~ true_state + utterance + goal, mss, mean)
ms$cih <- aggregate(goal_prob ~ true_state + utterance + goal, mss, ci.high)$goal_prob
ms$cil <- aggregate(goal_prob ~ true_state + utterance + goal, mss, ci.low)$goal_prob
```

## Expt 1 (most recent): goal attribution given utterance and heart-scale state

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 2.5}
ms$goal <- ordered(ms$goal, levels = c("honest", "nice", "mean"))
levels(ms$utterance) <- c("utterance: terrible", "utterance: bad","utterance: okay","utterance: good","utterance: amazing")

qplot(true_state, goal_prob, 
      colour = goal,
      data=ms) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~utterance) +
  xlab("state") +
  ylab("likelihood for goal") +
 # ggtitle("Experiment: goals given heart-state and utterance") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1))

# # only honesty goal
# qplot(true_state, goal_prob, 
#       colour = utterance,
#       data=subset(ms, goal=="honest")) + 
#   geom_line(aes(group=goal)) +
#   facet_grid(.~utterance) +
#   xlab("state (1=worst, 5=best)") +
#   ylab("likelihood for goal") +
#   ggtitle("honesty goal attribution given state and utterance (heart-scale)") +
#   geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1))

d3 <- d
ms3 <- ms
```


```{r warning=FALSE, message=FALSE, echo=FALSE}
# expt (1)
source("/Users/ericang/Documents/Research/polite_probmods/experiment/data_analysis/helper/useful.R")

raw.data.path <- "/Users/ericang/Documents/Research/polite_probmods/polite_adj/experiment/exp_versions/3_goals_v3/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```

## Expt 2: goal attribution given utterance and word-scale state

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 2.5}
rearr <- all.data %>%
  select(subid, num_range("answers.data.goal", 0:2)) %>%
  distinct(subid) %>%
  gather(goal_order, goal, num_range("answers.data.goal", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 18, 18))

rearr2 <- all.data %>%
  select(subid, num_range("answers.data.prediction", 0:1)) %>%
  distinct(subid) %>%
  gather(pred_order, prediction, num_range("answers.data.prediction", 0:1)) %>%
  mutate(pred_order = substr(pred_order, 24, 24))

d <- all.data %>%
  select(subid, answers.data.order, answers.data.knowledge, answers.data.domain, answers.data.state, answers.data.utterance, num_range("answers.data.predictedProb", 0:1), num_range("answers.data.inferredProb", 0:2)) %>%
  gather(pred_order, pred_prob, num_range("answers.data.predictedProb", 0:1)) %>%
  mutate(pred_order = substr(pred_order, 27, 27)) %>%
  gather(goal_order, goal_prob, num_range("answers.data.inferredProb", 0:2)) %>%
  mutate(goal_order = substr(goal_order, 26, 26))


detach(package:dplyr)
library(plyr)
d <- join(d, rearr)
d <- join(d, rearr2)

detach(package:plyr)
library(dplyr)
d <- d %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(true_state = answers.data.state) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(knowledge = answers.data.knowledge) %>%
  select(subid, trial, knowledge, true_state, utterance, goal, goal_prob, prediction, pred_prob)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$true_state <- as.factor(d$true_state) 
d$knowledge <- as.factor(d$knowledge)
d$utterance <- as.factor(d$utterance)
d$goal <- as.factor(d$goal)
d$prediction <- as.factor(d$prediction)
d$goal_prob <- as.numeric(d$goal_prob)
d$pred_prob <- as.numeric(d$pred_prob)

d$true_state <- ordered(d$true_state, levels = c("terrible", "bad", "okay", "good", "amazing"))
d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))

# goal_prob ~ true_state + utterance + goal + knowledge
mss <- d %>%
  group_by(true_state, utterance, goal, knowledge, subid) %>%
  summarize(
            goal_prob = mean(goal_prob, na.rm=TRUE)
          )
ms <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, mean)
ms$cih <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.high)$goal_prob
ms$cil <- aggregate(goal_prob ~ true_state + utterance + goal + knowledge, mss, ci.low)$goal_prob

ms$goal <- ordered(ms$goal, levels = c("honest", "nice", "mean"))
# levels(ms$true_state) <- c("state: terrible", "state: bad", "state: okay", "state: good", "state: amazing")
levels(ms$utterance) <- c("utterance: terrible", "utterance: bad", "utterance: okay", "utterance: good", "utterance: amazing")

# unknown
qplot(true_state, goal_prob, 
      colour = goal,
      data=subset(ms, knowledge == "unknown")) + 
  geom_line(aes(group=goal)) +
  facet_grid(.~utterance) +
  xlab("state") +
  ylab("likelihood for goal") +
  ggtitle("What was the speaker's goal? (with state-utt match)") +
  geom_errorbar(aes(ymin=goal_prob-cil,ymax=goal_prob+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

d1 <-d
ms1 <- subset(ms, knowledge == "unknown")
```


```{r warning=FALSE, message=FALSE, echo=FALSE}
# expt (2)
source("/Users/ericang/Documents/Research/polite_probmods/experiment/data_analysis/helper/useful.R")

raw.data.path <- "/Users/ericang/Documents/Research/polite_probmods/polite_adj/experiment/exp_versions/4_L2_J/production-results/"

## LOOP TO READ IN FILES
all.data <- data.frame()
files <- dir(raw.data.path,pattern="*.json")

for (file.name in files) {
  
  ## these are the two functions that are most meaningful
  json_file <- readLines(paste(raw.data.path,file.name,sep=""))
  json_file_str = paste(json_file, collapse = "")
  json_file_str = gsub(",}", "}", json_file_str)
  jso = jsonlite::fromJSON(json_file_str)
  jso$answers$data$people <- NULL
  jso1 <- data.frame(jso)
  jso1$subid <- substring(file.name, 1, 6)
  
  ## now here's where data get bound together
  all.data <- rbind(all.data, jso1)
}
```


## Expt 3: word (feeling) acceptance given heart-scale state

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 2.5}
d <- all.data %>%
  select(subid, answers.data.order, answers.data.domain, answers.data.state, answers.data.utterance, answers.data.judgment) %>%
  mutate(trial = answers.data.order) %>%
  mutate(item = answers.data.domain) %>%
  mutate(state = substr(answers.data.state, 6, 6)) %>%
  mutate(utterance = answers.data.utterance) %>%
  mutate(judgment = answers.data.judgment) %>%
  select(subid, trial, item, state, utterance, judgment)

d$subid <- as.factor(d$subid)
d$trial <- as.numeric(d$trial)
d$state <- as.factor(d$state) 
d$utterance <- as.factor(d$utterance)
d$judgment <- as.numeric(as.character(d$judgment))

d$utterance <- ordered(d$utterance, levels = c("terrible", "bad", "okay", "good", "amazing"))

# goal_prob ~ true_state + utterance + goal
mss <- d %>%
  group_by(state, utterance, subid) %>%
  summarize(
            judgment = mean(judgment, na.rm=TRUE)
          )
ms <- aggregate(judgment ~ state + utterance, mss, mean)
ms$cih <- aggregate(judgment ~ state + utterance, mss, ci.high)$judgment
ms$cil <- aggregate(judgment ~ state + utterance, mss, ci.low)$judgment

qplot(state, judgment, 
      colour = utterance,
      data=ms) + 
  geom_line(aes(group=utterance)) +
  facet_grid(.~utterance) +
  xlab("state") +
  ylab("judgment") +
  ggtitle("Semantic distribution of words (state of 1 = worst, 1-heart rating)") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1))

d2 <- d
ms2 <- ms
```

# Original vs. replication

Exclude expt3, which is about semantic distribution (should proximate honesty)

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 6}
ms1 <- ms1 %>%
  mutate(state = true_state) %>%
  mutate(judgment = goal_prob) %>%
  select(state, utterance, goal, judgment, cih, cil)
ms1$expt <- "word-state"
levels(ms1$state) <- c("1", "2", "3", "4", "5")
levels(ms1$utterance) <- c("terrible", "bad", "okay", "good", "amazing")

ms3 <- ms3 %>%
  mutate(state = true_state) %>%
  mutate(judgment = goal_prob) %>%
  select(state, utterance, goal, judgment, cih, cil)
ms3$expt <- "heart-state"

ms <- rbind(ms1, ms3)

qplot(utterance, judgment, 
      colour = expt,
      data=ms) + 
  geom_line(aes(group=expt)) +
  facet_grid(goal~state) +
  xlab("utterance") +
  ylab("judgment") +
  ggtitle("goals given state and utterance, facet by state") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1)) +
  theme(axis.text.x  = element_text(angle=45, vjust=0.5))

```

Differences:
1. people don't think it's less nice or more mean to say "good"" for 5-heart state compared to 4-heart state
2. saying "okay" is perceived to be less nice and more mean given state of 3, compared to an "okay" true state


# Comparing expts: honesty

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 2.5}
# rbind all three
ms1 <- ms1 %>%
  filter(goal =="honest") %>%
  select(state, utterance, judgment, cih, cil)
ms1$expt <- "honesty given word-state"

ms3 <- ms3 %>%
  filter(goal =="honest") %>%
  select(state, utterance, judgment, cih, cil)
ms3$expt <- "honesty given heart-state"

ms2$expt <- "word acceptance given heart-state"

ms <- rbind(ms1, ms2, ms3)
```


```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width = 10, fig.height = 2.5}
# separately showing the three expts
# honesty given heart state
qplot(state, judgment, 
      colour = utterance,
      data=ms3) + 
  geom_line(aes(group=utterance)) +
  facet_grid(.~utterance) +
  xlab("state (1=worst, 5=best)") +
  ylab("judgment") +
  ggtitle("honesty given heart state") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1))

# honesty given word state
qplot(state, judgment, 
      colour = utterance,
      data=ms1) + 
  geom_line(aes(group=utterance)) +
  facet_grid(.~utterance) +
  xlab("state (1=worst, 5=best)") +
  ylab("judgment") +
  ggtitle("honesty given word state") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1))

# word acceptance given heart state
qplot(state, judgment, 
      colour = utterance,
      data=ms2) + 
  geom_line(aes(group=utterance)) +
  facet_grid(.~utterance) +
  xlab("state (1=worst, 5=best)") +
  ylab("judgment") +
  ggtitle("word acceptance given heart state") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1))

# just expts using heart state
qplot(state, judgment, 
      colour = expt,
      data=subset(ms, expt!="honesty given word-state")) + 
  geom_line(aes(group=expt)) +
  facet_grid(.~utterance) +
  xlab("state (1=worst, 5=best)") +
  ylab("judgment") +
  ggtitle("acceptance/honesty rating given HEART-state and utterance") +
  geom_errorbar(aes(ymin=judgment-cil,ymax=judgment+cih,width=.1))
```

Honesty attribution with word-state looks more comparable to semantic distribution data?