---
title: "\"I won't lie to you, your talk wasn't amazing\": Modeling polite (in)direct remarks"
bibliography: politeness.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 

 \author{{\large \bf Erica J. Yoon}, {\large \bf Michael Henry Tessler}, {\large \bf Noah D. Goodman} \and {\large \bf Michael C. Frank}  \\
         \{ejyoon, mtessler, ngoodman, mcfrank\} @stanford.edu \\ 
         Department of Psychology, Stanford University}

abstract: 
    "Why do people speak politely? Previous work has suggested that people expect others to speak based on their desires to be transfer accurate information (epistemic goal) and to make the listener feel good (social goal), sometimes causing them to produce white lies [@yoon2016]. In the current work, we expand on this theory to consider another prominent case of polite speech: indirect remarks. We show that people expect a speaker to produce more of vague indirect remarks (e.g. \"It wasn't amazing\") when there is greater risk of face threat and the speaker wants to be considerate and informative at the same time. We also compare goal attributions to speakers who produce direct versus indirect remarks, and find that, when face threat risks are high, people attribute more extreme tradeoff decisions between the epistemic vs. social goal for direct remarks, but report greater balance between the two goals for indirect remarks. FIXME: These results align with our model predictions, demonstrating great generalizability of our model.
    "

keywords:
    "Politeness; computational modeling; communicative goals; pragmatics"
    
output: cogsci2016::cogsci_paper
---


```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Introduction

Language users hear and produce *polite speech* on a daily basis. But being polite conflicts with one important goal of cooperative communication: exchanging information efficiently and accurately [@Grice1975]. To be polite, people produce indirect requests that are much longer than simple imperatives ("It would be great if you could close that window" as opposed to "Close that window."), and tell white lies to make others feel good ("Your new dress is gorgeous!") Thus, speakers convey information inefficiently and risk losing accurate information (indirect remarks) or even intentionally convey wrong information (lies). If information transfer was the only currency in communication, a cooperative speaker would find polite utterances undesirable because they are potentially misleading. 

However, people do speak politely. Adults and even young children spontaneously produce requests in polite forms [@clark1980; @axia1985], and speakers use politeness strategies even while arguing, preventing unnecessary offense to their interactants [@holtgraves1997]. 

Do these facts about politeness imply that people are not cooperative communicators in the Gricean sense? @Brown1987 recast the notion of a *cooperative speaker* as one who has both an epistemic goal to improve the listener's knowledge state as well as a social goal to minimize any potential damage to the hearer's (and the speaker's own) self-image, which they called *face*. In their analysis, if the speaker's intended meaning contains no threat to the speaker or listener's face, then the speaker will choose to convey the meaning in an efficient manner, putting it *on the record*. As the degree of face-threat becomes more severe, however, a speaker will choose to be polite by producing more indirect utterances. 

Based on @Brown1987, in our previous work, we argued that language users think about polite speech as reflecting a tradeoff between information transfer and face-saving [@yoon2016]. When a speaker tries to save face, she hides or risks losing information in her intended message by making her utterance false or indirect to some degree. On the other hand, when a speaker prioritizes truthfulness and informativity, she may risk losing the listener's (or the speaker's own) face. We developed a novel computational model that captures the idea that cooperative speakers attempt to balance between the two goals: information transfer and face-saving. This model built on a recent formal framework for modeling pragmatic language understanding, the "rational speech act" (RSA)  model [@Frank2012; @Goodman2013]. We examined one specific phenomenon of polite speech, white lies (e.g. saying someone's performance was "okay" when in fact it was terrible), and predictions from our model were confirmed by empirical data on people's inferences about the relation between a speaker's goals (e.g. to be nice vs. honest), utterances ("It was okay") and the true states of the world (objectively how good was the addressee's piano recital performance). 

In this work, we expand on our previous work and examine another case study of polite speech phenomena: indirect remarks. Through indirect remarks, speakers try to convey a particular message in a more nuanced way. This is different from white lies, in that speaker's intentions are no longer hidden, but revealed with suboptimal efficiency. 

Why would people speak indirectly? For example, why would it be better to say "I would love another glass of wine, thanks." than "Pour me more wine"? The latter more clearly conveys the guest's intention for the waiter to pour more wine. But the former is less imposing on the waiter, and circumvents an impression that the speaker is in a position to give orders to the listener. Indeed, even when the implied meaning of the requests is the same, people prefer requests whose literal meanings ask for the listener's permission ("Could I ask you where Jordan Hall is?") to those with literal meanings that assume listener's obligation to respond ("Shouldn't you tell me where Jordan Hall is?") [@clark1980]. Indirect requests are complicated to manipulate for many reasons, for example due to various possible semantic forms of imperatives, and our proposed work focuses on *indirect remarks* as a simpler case study to begin with.

Indirect remarks may be used to reflect speaker's attempt to balance between two goals assumed by our model: informativity and face-saving. For example, if Ann hesitantly comments on a colleague's past presentation by saying "The presentation *wasn't amazing*..." she does not preclude the possibility that the presentation was bad, so the utterance is now not a downright lie; and Ann still tries to save Bob’s face by not explicitly saying that it was bad. Thus we hypothesize that, similar to white lies, indirect speech reflects speaker's desires to balance between the goal to be informative (convey information in the most direct manner possible) and the goal to save the listener's face (make the listener feel good, or at least avoid making them feel bad).

# Computational Model

In the current work, we build on our previous formal model that assumes speaker to choose utterances approximately optimally given a utility function, a standard assumption made in family of RSA models [@yoon2016; @Goodman2013]. We proposed that there are two utilities considered by the speaker. First, *epistemic utility* refers to the amount of information a *literal listener* would still not know about world state $s$ after hearing a speaker's utterance $w$ (*surprisal* that the speaker would want to minimize):
$$U_{epistemic}(w; s) = \ln(P_{L_0}(s \mid w))$$,
and the literal listener is a simple Bayesian agent that takes the utterance to be true:

$$
P_{L_0}(s \mid w)\propto {w}(s) \cdot P(s)
$$ 

Second, *social utility* is the expected utility of the state the listener would infer given the utterance $w$, which is related to the intrinsic value of the state from the listener's viewpoint :

$$
U_{social}(w; s) = \mathbb{E}_{P_{L_0}(s \mid w)}[V(s)],
$$
where $V$ is a value function that maps states to subjective utility values and thus captures the affective consequences for the listener of being in state $s$.

We defined the overall speaker utility to be a weighted combination of epistemic and social utilities:

$$
U(w;s;  \hat{\beta}) = \beta_{epistemic}\cdot U_{epistemic} + \beta_{social} \cdot U_{social}.
$$

The speaker chooses utterances $w$ softmax-optimally given the state $s$ and his goal weights $\hat{\beta}$: 
$$
P_{S_1}(w \mid s, \hat{\beta}) \propto \mathrm{exp}(\lambda \cdot \mathbb{E}[U(w; s;  \hat{\beta})])$$

The pragmatic listener, denoted $L_1$, infers the world state based on this speaker model. We will assume the listener does not know exactly how the speaker weights his competing goals, however.
We assume the pragmatic listener jointly infers the state $s$ and the utility weights of the speaker, $\beta_{epistemic}$ and $\beta_{social}$ [@GoodmanLassiter2015; @Kao2014]:

$$
P_{L_1}(s,  \hat{\beta} \mid w)\propto P_{S_1}(w \mid s,  \hat{\beta})\cdot P(s) \cdot P( \hat{\beta}) $$

Within our experimental domain, we assume there are five possible states of the world corresponding to the value placed on a particular referent (e.g. rating deserved by the presentation the speaker is commenting on): $S = \{s_{1}, ...,  s_{5}\}$. We further assume a uniform prior distribution over possible states of the world. The states have subjective numerical values $V(s_{i}) = \alpha \cdot i$, where $\alpha$ is a scaling parameter (later inferred from data).

The current work builds on our previous formal model by adding simple but key components to predict indirect remark production and comprehension. First, there had previously been five possible utterances: {It was *terrible*, *bad*, *okay*, *good*, and *amazing*}, all direct assertions of specific states (e.g. "It was amazing" would be true for the state of 5 but untrue for the states of 1 or 2). To probe people's inferences about indirect remarks, we added five utterances to the set: {It *wasn't* terrible, bad, okay, good, and amazing}. These utterances indirectly address the referent by negating certain state. Second, the cost of longer utterances (it is more costly to say "It wasn't terrible" than "It was amazing" due to inclusion of negation; @GoodmanLassiter2015) was accounted for by assigning greater cost to indirect remarks with negation (cost = 2) than direct remarks with no negation (cost = 1), which affected the likelihood of production of utterances *a priori*. We implemented this model using the probabilisitic programming language WebPPL [@dippl] and a complete implementation can be found at \url{FIXME}.

To confirm the predictions of our new model, we first measure the literal semantics in Experiment 1, then use these to predict people's responses in two subsequent experiments. In Experiment 2, we examine people's expectations for speaker production of utterances $u$ given a state and a speaker's goal. In Experiment 3, we investigate listeners' inferences about speakers' goals given an utterance and a state. Then we compare the behavioral results to our model predictions, that people's inferences are based on a model of a speaker with two utilities: epistemic and social. 

# Experiment 1: Literal semantics

Experiment 1 probed judgments of literal meanings of the target words assumed by our model and used in all our experiments. These judgments will be used as the expected literal meanings in our model.

```{r expt1_screen, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Example of a trial in Experiment 1."}
img <- png::readPNG("figs/expt1_screen.png")
grid::grid.raster(img)
```

## Method 

### Participants 

25 participants with IP addresses in the United States were recruited on Amazon's Mechanical Turk.

### Stimuli and Design

We used 13 different context items that were previously used in @yoon2016, in which someone evaluated a performance of some kind (e.g. presentation). For example, in one of the contexts, Bob saw a presentation, and Bob's feelings toward Ann's cake (*true state*) were shown on a scale out of five hearts (e.g. two out of five hearts filled in red color). The question of interest was "Do you think Bob thought the presentation was / wasn't X?" where X could be one of five possible words: *terrible*, *bad*, *okay*, *good*, and *amazing*, giving rise to ten different possible utterances (with negation or no negation). Each participant read 50 scenarios, depicting every possible combination of 5 true states and 10 utterances. The order of context items was randomized, and there were a maximum of four repeats of each context item per participant.

### Procedure

Participants read scenarios and indicated their answer to each question by answering "No" or "Yes" (see Figure 1 for a screenshot of an example trial). The experiment can be viewed at: \url{https://langcog.stanford.edu/expts/EJY/polgrice/negimp_prior_v1/negimp_prior.html}.

```{r expt1_results, fig.env = "figure*", fig.pos = "b", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Results from Experiment 1."}
img <- png::readPNG("figs/expt1_results.png")
grid::grid.raster(img)
```

### Results 

For this and all subsequent experiments, we analyze the data by collapsing across context items. Meanings of the words as judged by participants were as one would expect (see Figure 2). For utterances without negation (e.g. "It was terrible"), proportion of acceptances for a word given the true state peaked where the degree of positivity, neutrality and negativity of the state matched that of the word, replicating literal semantics reported in @yoon2016. For utterances with negation (e.g. "It wasn't terrible"), proportion of acceptances were inverses of acceptances for non-negated words. The fraction of participants that endorsed utterance $w$ for state $s$ will be used as the literal meaning ${w}(s)$ in Eq. 1.

# Experiment 2: Speaker production

In Experiment 2, we examined people's predictions for the most likely utterance produced by the speaker ($u$), given a description of the true state of the world (e.g. the speaker felt that a poem deserved 2 out of 5 hearts) and the speaker's goals (e.g. the speaker wanted to make the listener feel good). Critically, the contexts indicated face threats toward the listener, as the speaker's utterance was an evaluation of the listener's performance. We hypothesized that the tradeoff in the speaker's intention to be informative versus to save the listener's face would lead to greater use of vague, indirect remarks, especially when the performance was poor.

## Method

### Participants

202 participants with IP addresses in the United States were recruited on Amazon's Mechanical Turk.

### Stimuli and Design

We designed scenarios in which a person (e.g. Ann) gave some performance and asked for another person (e.g. Bob)'s opinion on the performance. The same context items and true states as Experiment 1 were used. Additionally, we provided information on the speaker Bob's goal (*to make Ann feel good*, or *to give as accurate and informative feedback as possible*, or both) and the true state, or how Bob actually felt Ann's performance (e.g. 2 out of 5 hearts), Then we asked participants to predict what Bob would say, out of 10 possible utterances ("It was terrible", "It was bad" ... "It wasn't good", "It wasn't amazing"). Each participant read 15 scenarios, depicting every possible combination of 3 goals and 5 states. The order of context items was randomized, and there were a maximum of two repeats of each context item per participant.

### Procedure

Participants read each scenario followed by a question that read, "If Bob wanted *to make Ann feel good* (or *to give accurate and informative feedback*, or *BOTH make Sarah feel good AND give accurate and informative feedback*), what would Bob be most likely to say?" Participants indicated their answer by choosing one of the options on the two dropdown menus, side-by-side, one for choosing between *was* vs. *wasn't* and the other for choosing among *terrible*, *bad*, *okay*, *good*, and *amazing* (see Figure 3).
The experiment can be viewed at: \url{https://langcog.stanford.edu/expts/EJY/polgrice/speaker_production_dropdown_v2/speaker.html}.


## Behavioral results

```{r expt2_screen, fig.env = "figure", fig.pos = "b", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Example of a trial in Experiment 2."}
img <- png::readPNG("figs/expt2_screen.png")
grid::grid.raster(img)
```

People's predictions for speaker's utterances given varied depending on speaker's goals, and these differences wre especially pronounced for worse true states (Figure 4). 
For good states (4 and 5 hearts), positive direct remarks were judged to be the most likely utterances across all three goal conditions. For less-than-perfect, but still decent states, there was a greater degree of expectation of white lies (e.g. "It was amazing" for 4 hearts) given social goal. 
Our hypotheses were borne out for bad states (1 and 2 hearts): there were more instances of expected indirect remarks overall across all goal conditions given bad states. Critically, speakers with both goals to be informative and socially considerate were predicted to produce more indirect than direct remarks, unlike the other two goal conditions (Figure 5). Thus, these results indicated that a speaker who considers both informative and social goals, and thus is in want of a compromise between the two, is expected to produce relatively more indirect remarks.

```{r expt2_results, fig.env = "figure*", fig.pos = "b", fig.width=8, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Results from Experiment 2."}
img <- png::readPNG("figs/expt2_results.png")
grid::grid.raster(img)
```

```{r expt2_results2, fig.env = "figure", fig.pos = "t", fig.width=4, fig.height=8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Results from Experiment 2."}
img <- png::readPNG("figs/expt2_results2.png")
grid::grid.raster(img)
```
## Model predictions

### Model fitting

### Results

# Experiment 3: Goal inference

## Method

### Participants

60 participants with IP addresses in the United States were recruited on Amazon's Mechanical Turk.

### Stimuli and Design

We presented the same context items and true states (i.e. how Bob actually felt towards Ann's performance) as Experiment 2, but instead of goals we provided Bob's utterances ("It wasn't amazing"). Then we asked participants to infer the likelihood of Bob's goals to *to make Ann feel good*, or *to give accurate and informative feedback*. Each participant was randomly assigned to one of two conditions to read 25 scenarios, depicting half (counterbalanced) of all possible combinations of 5 true states and 10 utterances. The order of context items was randomized, and there were a maximum of two repeats of each context item per participant.

### Procedure

Participants read each scenario followed by a question that read, "Based on what Bob said, how likely do you think that Bob's goal was to: *to make Ann feel good*; *to give accurate and informative feedback*" with the two goals placed in a random order next to two slider bars, on which the participant could indicate each goal's likelihood. 
The experiment can be viewed at: \url{https://langcog.stanford.edu/expts/EJY/polgrice/L2_G_Neg/polgrice_L2_G.html}.

## Behavioral results

```{r expt3_results, fig.env = "figure*", fig.pos = "b", fig.width=8, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Results from Experiment 2."}
img <- png::readPNG("figs/expt3_results.png")
grid::grid.raster(img)
```

Participants rated speaker's goals differentially depending on the true state and utterance (see Figure 6).
Results for direct remarks with no negation replicated our previous findings: 
Goal to be informative was rated highest when the true state was most consistent with the literal meanings. 
As direct remarks became more positive, participants increased in their attribution of speaker's goal to make the listener feel good. 

Comparison of direct versus indirect remarks revealed an interesting asymmetry. 
For positive states (4 and 5 hearts), both informative and social goal attributions increased with more positive valence of the direct remarks, whereas both goal attributions mostly stayed low below chance level across all indirect remarks.
For negative states (1 and 2 hearts), interesting similarities and differences between the two types of utterances were observed. Similarly between the two, there was a crossover in goal attribution, caused by tradeoff between utterance-meaning match (informative goal) and face-saving (social goal). 
This tradeoff however is much more exaggerated in direct than indirect remarks: saying "It was amazing" given a state of 1 heart blatantly prioritizes social goal and forgos informative goal. 
On the other hand, indirect remarks present a more balanced decision: saying "It wasn't amazing"  moderately satisfies both goals. Again, this finding confirms that a speaker who seeks a compromise between the goals to be informative and to save face produces more indirect remarks, and this desire then to be recognized by the listener. 

## Model predictions

### Model fitting

### Results

# Discussion

In this work, we showed that our formal model with two speaker utilities (epistemic and social) can be used to not only explain white lie understanding but also indirect remark production and comprehension. As we predicted, speakers were expected to produce more indirect remarks when the listener's performance of concern was poorer, and thus the listener's face was under threat, and when the speakers wanted a compromise between the two conflicting goals (Experiment 2). Consistent with this, participants inferred a greater balance of the two goals in production of indirect remarks given bad true states. 

An important contribution of this work is in showing the generalizability of our formal model in two ways: extending to other kinds of polite speech (white lies and indirect remarks), and predicting comprehension as well as production of polite speech. 

FIXME: some discussion about other approaches to indirect speech (e.g. Danescu-Niculescu-Mizil's work on polite requests)

# Acknowledgements

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
